### 2025-01-13 Figures/PaperFiguresextended.Rmd
Added code to create figures

---
title: "Figures"
author: "LD"
date: '2023-06-06'
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---
```{r setup}
library(Seurat)
library(dplyr)
library(tibble)
library(stringr)
library(ggplot2)
library(cowplot)
library(ggbreak) 
library(tidyr)
library(patchwork)
library(ComplexUpset)
library(ggrepel)
library(Polychrome)
library(ComplexHeatmap)
library(gplots)
library(RColorBrewer)
library(randomcoloR)

dir_out <- "/home/common/data/output/projects/ENDO/E044/A033/"
dir.create(dir_out, recursive = TRUE)
setwd(dir_out)
```

In script
Replaced ProlifPeriSecFinal with MenstrualCyclePhase_main

```{r save downsampled data}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")

# Downsample the number of cells per identity class
Idents(data) <- data$sample
data_ds <- subset(x = data, downsample = 200)
saveRDS(data_ds, "/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")

#Some plotting
DimPlot(data, group.by = "AnnotationMain", raster = FALSE)
DimPlot(data, group.by = "Symphony_Refined_A029", raster = FALSE)
DimPlot(data, group.by = "Symphony_Refined_A029_United", raster = FALSE)

#Save object subsetted for ScaiNet samples
ScaiNETsamples <- read.csv("/home/common/data/output/projects/ENDO/E041/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv")$names
Idents(data) <- data$sample
subset <- subset(data, idents = ScaiNETsamples)
rm(data)
gc()
meta <- subset@meta.data
saveRDS(subset, "/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_ScaiNETsamples.rds")
rm(subset)
gc()

meta$AnnotationMain

```

Remove? TODO
```{r order of samples by menstrual cycle phase and pseudotime}
meta_final <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")

MenstrualCyclePhase_sampleOrder <- meta_final %>%
  arrange(MenstrualCyclePhase,epithelial.pseudotime) %>%
  distinct(sample,epithelial.pseudotime, MenstrualCyclePhase,CycleDay) %>%
  pull(sample)
saveRDS(MenstrualCyclePhase_sampleOrder, paste0(dir_out,"sampleorder_MenstrualCyclePhase_greenCirclemedian.RDS"))
```

Add additional data to the Seurat metadata and reformat
```{r plotting of cell types per sample}
#Set directory links and create directories
dir_data <- "L:/GitHubDesktop/endometriosis_endometrium_scRNA_atlas/_Data/"
dir_out <- "L:/GitHubDesktop/endometriosis_endometrium_scRNA_atlas/Figures_out/"
if (!dir.exists(dir_out)) {
  dir.create(dir_out)
}

#Load data
#A003A033 <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_metadata_ptime_snn_Final.rds")
meta <- readRDS(paste0(dir_data, "EndoAtlas_GEO_meta_complete.rds"))

##Modify menstrual cycle information
# Create column MenstrualCyclePhase_main based on MenstrualCyclePhase by merging the secretory sub-phases from 
meta$MenstrualCyclePhase_main <- as.character(meta$MenstrualCyclePhase)
meta$MenstrualCyclePhase_main[meta$MenstrualCyclePhase_main %in% c("secretory_early", "secretory_mid", "secretory_late")] <- "secretory"
#Factor samples by MenstrualCyclePhase_main for plotting
meta$MenstrualCyclePhase_main <- factor(meta$MenstrualCyclePhase_main, levels = c("proliferative","periovulatory","secretory"))

##Add additional metadata from manuscript Supplementary Table 1
# Read the tab-delimited text file
SupplementaryTable1 <- read.delim(paste0(dir_data, "SupplementaryTable1.txt"), header = TRUE, stringsAsFactors = FALSE)
# Perform a left join of the `meta` dataframe with `SupplementaryTable1`
meta <- dplyr::left_join(
  meta,
  SupplementaryTable1 %>% 
    dplyr::select(
      Patient.sample.ID,
      epithelial.pseudotime,
      sequencing.batch,
      Minor.exclusion.criteria.logical,
      Menstrual.cycle.phase.marker.analysis,
      Cell.type.frequency.analysis..Figure.2e.f.,
      DEG.analysis..Figure.3a.b.,
      Ligand.receptor.analysis..Figure.3c.h.
    ),
  by = c("sample" = "Patient.sample.ID")
)

#Double-check, TODO
table(meta$sample, meta$Minor.exclusion.criteria.logical)
prolifsamples <- meta%>%
  dplyr::filter(MenstrualCyclePhase_main == "proliferative") %>%
  select(sample) %>%
  distinct() %>%
  arrange(sample) %>%
  pull(sample)
DEGsamples <- unique(read.csv("/home/common/data/output/projects/ENDO/E044/A030/major_celltypes_Tan/celltype_freq_sample.csv")$Var2)
DEGsamples <- DEGsamples[order(DEGsamples)]
setdiff(prolifsamples,DEGsamples) #MinorExclusion samples overlap with samples excluded in DEG analysis, sample87 is late secretory

##Determine order of samples by MenstrualCyclePhase and epithelial.pseudotime
MenstrualCyclePhase_sampleOrder <- meta%>%
  select(sample, MenstrualCyclePhase,epithelial.pseudotime) %>%
  distinct() %>%
  arrange(MenstrualCyclePhase, epithelial.pseudotime) %>%
  mutate(MenstrualCyclePhase_sampleOrder = row_number()) %>%
  ungroup()
#Factor samples by MenstrualCyclePhase and epithelial.pseudotime for plotting
meta$sample <- factor(meta$sample, levels = c(MenstrualCyclePhase_sampleOrder$sample))

#Save color pals
qual_col_pals <- brewer.pal.info[brewer.pal.info$category == "qual", ]
col_vector <- unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

Supplementary Data Figure 4 a
```{r plotting of cell types per sample}
#########################################################################
# bar plot of the number of AnnotationRefined cells in each sample
df <- meta%>%
  filter(!is.na(AnnotationMain)) %>%
  group_by(sample, AnnotationMain, AnnotationRefined, EndometriosisGrade, MenstrualCyclePhase, epithelial.pseudotime, EndometriosisStatus) %>%
  summarise(row_count = n()) %>%
  ungroup()

#Bar plot
#####use this for paper supplement
#Bar plot split by main5 
#MenstrualCyclePhase_sampleOrder <- readRDS(paste0(dir_out,"sampleorder_MenstrualCyclePhase_greenCirclemedian.RDS"))
#df$sample <- factor(df$sample, levels = MenstrualCyclePhase_sampleOrder)
P63 = createPalette(63, c("#ff0000", "#00ff00", "#0000ff")) 
swatch(P63)
names(P63) <- NULL
#Make barplots with different seeds
seed <- 1 #find in dir.create(paste0(dir_out, "/seedopt/"))
set.seed(seed)  
p <- ggplot(df, aes(x = sample, y = row_count, fill = AnnotationRefined)) +
  geom_bar(stat = 'identity', position = 'fill') +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = sample(P63)) +
  xlab('Sample ID') +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "lightgray", size = 0.5)
  ) +
  facet_grid(AnnotationMain ~ ., scales = "free_y")
ggsave(paste0(dir_out, "/seedopt/Barplot_sample_AnnotationRefined_main5_", seed, ".pdf"),
       plot = p,
       width = 15, height = 6)

##Consider leaving out
a1 <- ggplot(df %>% select(sample, EndometriosisStatus) %>% distinct (), aes(x = sample, y = EndometriosisStatus, fill = EndometriosisStatus)) +
  geom_bar(stat = 'identity', position = 'fill') +
  xlab('Sample ID') +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "lightgray", size = 0.5)
  ) 
a2 <- ggplot(df %>% select(sample, MenstrualCyclePhase) %>% distinct (), aes(x = sample, y = MenstrualCyclePhase, fill = MenstrualCyclePhase)) +
  geom_bar(stat = 'identity', position = 'fill') +
  xlab('Sample ID') +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(colour = "lightgray", size = 0.5)
  )
a1 + a2 + plot_layout(nrow = 2)
ggsave(paste0(dir_out, "Barplot_sample_AnnotationRefined_main5_annotation.pdf"),
  plot = last_plot(),
  width = 15, height = 6)

```

Figure 2 e
```{r plotting of cell types per sample}
##########################################################################
#BoxPlots sample, AnnotationMain, EndometriosisGrade, MenstrualCyclePhase_main
df2 <- meta%>%
  filter(Minor.exclusion.criteria.logical == "FALSE") %>%
  group_by(sample, AnnotationMain, EndometriosisGrade, MenstrualCyclePhase_main, EndometriosisStatus) %>% #no sample, AnnotationMain combinations are dropped
  summarise(row_count = n()) %>%
  ungroup() %>%
  filter(!is.na(AnnotationMain)) %>%
  group_by(sample) %>%
  mutate(AnnotationMain_perc = row_count / sum(row_count) * 100) %>%
  ungroup()

#BoxPlots percentage, EndometriosisStatus #for publication
ggplot(df2, aes(x = MenstrualCyclePhase_main, y = AnnotationMain_perc, fill = EndometriosisStatus)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(colour = EndometriosisStatus),
             colour="black",pch=21, 
             position = position_jitterdodge()) +
  RotatedAxis() +
  xlab('Sample ID') +
  facet_wrap(~AnnotationMain,scales = "free", ncol = 6)
ggsave(paste0(dir_out, "Boxplot_MenstrualCyclePhase_main_main5_sample_EndometriosisStatus_perc_publication.pdf"),
  plot = last_plot(),
  width = 12, height = 4.4)

###Significance testing CTL vs Endo
# Perform t-test CTL vs Endo for each cell type grouped by AnnotationRefined, MenstrualCyclePhase_main
df2_ttest_EndometriosisStatus <- df2 %>%
  group_by(AnnotationMain, MenstrualCyclePhase_main) %>%
  summarise(p.value = t.test(AnnotationMain_perc ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#epithelial, proliferative just not significant (p.value = 0.0898)

###Significance testing prolif vs periov vs sec, CTL + Endo
## Perform t-test Prolif vs Peri for each cell type
df2_ttest_ProlifPeri <- df2 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("proliferative","periovulatory")) %>%
  group_by(AnnotationMain) %>%
  summarise(p.value = t.test(AnnotationMain_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
#none significant
## Perform t-test Prolif vs Sec for each cell type
df2_ttest_ProlifSec <- df2 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("proliferative","secretory")) %>%
  group_by(AnnotationMain) %>%
  summarise(p.value = t.test(AnnotationMain_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
#epi (0.00977) and str (0.0255) significant

## Perform t-test periovulatory vs sec for each cell type
df2_ttest_PeriSec <- df2 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("periovulatory","secretory")) %>%
  group_by(AnnotationMain) %>%
  summarise(p.value = t.test(AnnotationMain_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
#epi (0.00794) and str (0.0137) significant
```

Supplementary Data Figure 4b, delete?
```{r plotting of cell types per sample}
##########################################################################################
##Plot differences for DEG analysis samples
#Extract sample names of samples used for DEG analysis
#DEGsamples <- unique(read.csv("/home/common/data/output/projects/ENDO/E044/A030/major_celltypes_Tan/celltype_freq_sample.csv")$Var2)

DEGsamples <- SupplementaryTable1 %>%
  filter(DEG.analysis..Figure.3a.b. == TRUE) %>%
  pull(Patient.sample.ID)

#Make df 
#Introduce 0 for the ones with 0 count!
df3 <- meta%>%
  group_by(sample, AnnotationRefined) %>%
  summarise(AnnotationRefined_counts = n()) %>%
  ungroup() %>%
  complete(sample, AnnotationRefined, fill = list(AnnotationRefined_counts = 0)) %>%
  filter(!is.na(AnnotationRefined)) %>%
  group_by(sample) %>%
  mutate(AnnotationRefined_perc = AnnotationRefined_counts / sum(AnnotationRefined_counts) * 100) %>%
  ungroup()  %>% #60 samples * 63 cell types = 3780 rows, correct
  filter(sample %in% DEGsamples) #23 samples * 63 cell types = 1449 rows, correct
df3 <- left_join(df3, dplyr::distinct(meta[,c("sample","EndometriosisStatus")]), by = "sample")

# Perform t-test for each cell type
df3_ttest <- df3 %>%
  group_by(AnnotationRefined) %>%
  summarise(p.value = t.test(AnnotationRefined_perc ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#No significant difference between the EndometriosisStatus in any cell type

###Plot
#BoxPlots percentage, EndometriosisStatus, percentage
ggplot(df3, aes(x = AnnotationRefined, y = AnnotationRefined_perc, fill = EndometriosisStatus)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(colour = EndometriosisStatus),
             colour="black",pch=21, 
             position = position_jitterdodge(jitter.width = 0.2)) +
  facet_wrap(~AnnotationRefined, scales = "free", ncol = 7)
ggsave(paste0(dir_out, "Boxplot_DEGsamples_SymphonyRefinedFinal_EndometriosisStatus_perc.pdf"),
  plot = last_plot(),
  width = 17, height = 12)
```

Supplementary Data Figure 4b
```{r plotting of cell types per sample}
##########################################################################################
##Plot differences for samples without minor exclusion criteria

#Make df 
#Introduce 0 for the ones with 0 count!
df4 <- meta%>%
  filter(Minor.exclusion.criteria.logical == "FALSE") %>%
  group_by(sample, AnnotationRefined) %>%
  summarise(AnnotationRefined_counts = n()) %>%
  ungroup() %>%
  complete(sample, AnnotationRefined, fill = list(AnnotationRefined_counts = 0)) %>%
  filter(!is.na(AnnotationRefined)) %>%
  group_by(sample) %>%
  mutate(AnnotationRefined_perc = AnnotationRefined_counts / sum(AnnotationRefined_counts) * 100) %>%
  ungroup() %>% #60 samples * 63 cell types = 3780 rows, correct
  filter(!sample %in% c("sample81", "sample87", "sample108", "sample115", "sample119", "sample121"))  #54 samples * 63 cell types = 3402 rows, correct
df4 <- left_join(df4, dplyr::distinct(meta[,c("sample","EndometriosisStatus","MenstrualCyclePhase_main")]), by = "sample")
write.csv(df4, paste0(dir_out,"df4_AnnotationRefined_frequency.csv"))

# Perform t-test CTL vs Endo for each cell type grouped by AnnotationRefined, MenstrualCyclePhase_main
df4_ttest_EndometriosisStatus <- df4 %>%
  group_by(AnnotationRefined, MenstrualCyclePhase_main) %>%
  summarise(p.value = t.test(AnnotationRefined_perc ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#Prv_VSMC secretory, secretory is the only significant one (pval 0.009075467)!!
df4_ttest_EndometriosisStatus$p.value_adjBH <- p.adjust(df4_ttest_EndometriosisStatus$p.value, method = "BH", n = length(df4_ttest_EndometriosisStatus$p.value))
#After correction for multiple testing Prv_VSMC secretory is not significant anymore

df4_ttest_EndometriosisStatus_secretory <- df4_ttest_EndometriosisStatus %>%
                      filter(MenstrualCyclePhase_main == "secretory")
df4_ttest_EndometriosisStatus_secretory$p.value_adjBH <- p.adjust(df4_ttest_EndometriosisStatus_secretory$p.value, method = "BH", n = length(df4_ttest_EndometriosisStatus_secretory$p.value))
#Even when subsetting secretory phase, after correction for multiple testing Prv_VSMC secretory is not significant anymore

## Perform t-test Prolif vs Peri for each cell type
df4_ttest_ProlifPeri <- df4 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("proliferative","periovulatory")) %>%
  group_by(AnnotationRefined) %>%
  summarise(p.value = t.test(AnnotationRefined_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
#Few significant
## Perform t-test Prolif vs Sec for each cell type
df4_ttest_ProlifSec <- df4 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("proliferative","secretory")) %>%
  group_by(AnnotationRefined) %>%
  summarise(p.value = t.test(AnnotationRefined_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
write.csv(df4_ttest_ProlifSec, paste0(dir_out,"df4_ttest_ProlifSec.csv"))
df4_ttest_ProlifSec <- read.csv(paste0(dir_out,"df4_ttest_ProlifSec.csv"))
df4_ttest_ProlifSec$p.value_adjBH <- p.adjust(df4_ttest_ProlifSec$p.value, method = "BH", n = length(df4_ttest_ProlifSec$p.value))
write.csv(df4_ttest_ProlifSec, paste0(dir_out,"df4_ttest_ProlifSec_adjpval.csv"))

#About half are significant, show in paper supplement
## Perform t-test periovulatory vs sec for each cell type
df4_ttest_PeriSec <- df4 %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("periovulatory","secretory")) %>%
  group_by(AnnotationRefined) %>%
  summarise(p.value = t.test(AnnotationRefined_perc ~ MenstrualCyclePhase_main)$p.value) %>%
  ungroup()
#Less than half are significant, don't show in paper supplement

#BoxPlots percentage, EndometriosisStatus, percentage ###for paper Extended Data Fig. S4
ggplot(df4, aes(x = MenstrualCyclePhase_main, y = AnnotationRefined_perc, fill = EndometriosisStatus)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(colour = EndometriosisStatus),
             colour="black",pch=21, 
             position = position_jitterdodge(jitter.width = 0.2)) +
  facet_wrap(~AnnotationRefined, scales = "free", ncol = 7) +
  theme_bw() +
  expand_limits(y = 0)

ggsave(paste0(dir_out, "Boxplot_DEGsamples_SymphonyRefinedFinal_EndometriosisStatus_perc.pdf"),
  plot = last_plot(),
  width = 20, height = 12)
```

Figure 2f
```{r plotting of cell types per sample}
#Early and Mid Secretory Prv_VSMCs 
df4 <- meta%>%
  filter(Minor.exclusion.criteria.logical == "FALSE") %>%
  group_by(sample, AnnotationRefined) %>%
  summarise(AnnotationRefined_counts = n()) %>%
  ungroup() %>%
  complete(sample, AnnotationRefined, fill = list(AnnotationRefined_counts = 0)) %>%
  filter(!is.na(AnnotationRefined)) %>%
  group_by(sample) %>%
  mutate(AnnotationRefined_perc = AnnotationRefined_counts / sum(AnnotationRefined_counts) * 100) %>%
  ungroup() %>% #60 samples * 63 cell types = 3780 rows, correct
  filter(!sample %in% c("sample81", "sample87", "sample108", "sample115", "sample119", "sample121"))  #54 samples * 63 cell types = 3402 rows, correct
df4 <- left_join(df4, dplyr::distinct(meta[,c("sample","EndometriosisStatus","MenstrualCyclePhase_main","MenstrualCyclePhase")]), by = "sample")

secEarylMid_Prv_VSMC_MenstrualCyclePhase_main <- df4 %>%
  filter(AnnotationRefined %in% c("Prv_VSMC secretory", "Prv_VSMC secretory_late")) %>% #same result with and without "Prv_VSMC secretory_late"
  dplyr::filter(MenstrualCyclePhase %in% c("proliferative","periovulatory","secretory_mid","secretory_early")) %>%
  group_by(sample, EndometriosisStatus, MenstrualCyclePhase_main) %>%
  summarize(
    total_counts_MenstrualCyclePhase_main = sum(AnnotationRefined_counts, na.rm = TRUE),
    total_perc_MenstrualCyclePhase_main = sum(AnnotationRefined_perc, na.rm = TRUE)
  ) %>%
  ungroup()

ggplot(secEarylMid_Prv_VSMC_MenstrualCyclePhase_main, aes(x = MenstrualCyclePhase_main, y = total_perc_MenstrualCyclePhase_main, fill = EndometriosisStatus)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(colour = EndometriosisStatus),
             colour="black",pch=21, 
             position = position_jitterdodge(jitter.width = 0.2))
ggsave(paste0(dir_out, "Boxplot_sec_Prv_VSMC_NOsecretory_late_total_perc_MenstrualCyclePhase_main_EndometriosisStatus.pdf"),
  plot = last_plot(),
  width = 17, height = 12)

secEarylMid_Prv_VSMC_MenstrualCyclePhase_main_pval <- secEarylMid_Prv_VSMC_MenstrualCyclePhase_main %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("secretory")) %>% 
  summarise(p.value = t.test(total_perc_MenstrualCyclePhase_main ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#significant, p.value 0.0261 -> paper

secEarylMid_Prv_VSMC_MenstrualCyclePhase_main_pval_prolif <- secEarylMid_Prv_VSMC_MenstrualCyclePhase_main %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("proliferative")) %>% 
  summarise(p.value = t.test(total_perc_MenstrualCyclePhase_main ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#significant, p.value NA

secEarylMid_Prv_VSMC_MenstrualCyclePhase_main_pval_peri <- secEarylMid_Prv_VSMC_MenstrualCyclePhase_main %>%
  dplyr::filter(MenstrualCyclePhase_main %in% c("periovulatory")) %>% 
  summarise(p.value = t.test(total_perc_MenstrualCyclePhase_main ~ EndometriosisStatus)$p.value) %>%
  ungroup()
#significant, p.value 0.7639853

p.adjust(c(0.0261), method = "BH", n = length(3)) #Value does not change
p.adjust(c(0.7639853), method = "BH", n = length(3)) #Value does not change

#Calculate fold change ENDO vs nonENDO secretory
# Assuming secEarylMid_Prv_VSMC_MenstrualCyclePhase_main is your dataframe

library(dplyr)

# Filter the dataframe for 'secretory' in MenstrualCyclePhase_main and then group by EndometriosisStatus
mean_total_perc <- secEarylMid_Prv_VSMC_MenstrualCyclePhase_main %>%
  filter(MenstrualCyclePhase_main == "secretory") %>%
  group_by(EndometriosisStatus) %>%
  summarize(mean_total_perc = mean(total_perc_MenstrualCyclePhase_main))
```

Figure 1d
```{r plotting of cell types per sample}
p5 <- ggplot(df5, aes(x = AnnotationMain, y = row_count)) +
  geom_col() +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip()
p5
ggsave(paste0(dir_out, "Barplot_EndometriosisGrade_main5.pdf"),
  plot = last_plot(),
  width = 15, height = 6)

#main5 boxplot 
df6 <- meta%>%
  group_by(sample, AnnotationMain, EndometriosisGrade) %>%
  summarise(row_count = n()) %>%
  ungroup()
p6 <- ggplot(df6, aes(x = AnnotationMain, y = row_count)) +
  geom_boxplot() +
  #geom_jitter(position = position_jitter(width = 0.2, height = 0), size = 0.5) +
  scale_y_log10() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p6
ggsave(paste0(dir_out, "Boxplot_EndometriosisGrade_main5_colEndometriosisGrade.pdf"),
         plot = last_plot(),
         width = 15, height = 6)

# bar plot of the number of AnnotationRefined cells in each sample
df7 <- meta%>%
  filter(!is.na(AnnotationMain)) %>%
  group_by(AnnotationMain, EndometriosisGrade) %>%
  summarise(row_count = n()) %>%
  ungroup() 
ggplot(df7, aes(x = AnnotationMain, y = row_count, fill = EndometriosisGrade)) +
  geom_bar(stat = 'identity', position = 'fill') +
  scale_y_continuous(expand = c(0, 0)) +
  coord_flip()
ggsave(paste0(dir_out, "Barplot_AnnotationMain_EndometriosisGrade.pdf"),
  plot = last_plot(),
  width = 6, height = 4)

```

UMAPs

epithelial UMAP
2b,2c,3b,3e,

atlas UMAP
S3f, S7a

Figure 1c
```{r}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")
A003A033 <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")
table(A003A033$AnnotationRefined)
data <- AddMetaData(data,A003A033)
E031A004meta <- read.csv("/home/common/data/output/projects/ENDO/E044/A033/E031A004_proliferative_SCT_percent.mt_sample_integration_batch_hvg_annotated_meta.csv")
E031A004meta$cell_id <- paste(E031A004meta$sample, E031A004meta$cell.names, sep = "_")
rownames(E031A004meta) <- E031A004meta$cell_id 
data <- AddMetaData(data, E031A004meta[, "Annotation", drop = FALSE])

order <- A003A033 %>%
  select(AnnotationMain, AnnotationRefined) %>%
  filter(!is.na(AnnotationMain)) %>%
  distinct() %>%
  arrange(AnnotationRefined) %>%
  mutate(AnnotationMain = factor(AnnotationMain, levels = c( "endothelial","stromal","lymphocyte","epithelial","myeloid"))) %>%
  arrange(AnnotationMain)
data$AnnotationRefined <- factor(x = data$AnnotationRefined, levels = order$AnnotationRefined)
p24 <- DimPlot(data, reduction = "umap", group.by = "AnnotationRefined",  raster = FALSE)
p24

```

Figure 2b
```{r}
#Figure 2b entire atlas
Fig2b1 <- DimPlot(data, reduction = "umap", group.by = "MenstrualCyclePhase", raster = FALSE)
Fig2b1
ggsave(filename = paste0(dir_out, 'Fig2b1.pdf'),
       plot = Fig2b1,
       width =20, height=9)

#Figure 2b epithelial cells
Fig2b2 <- DimPlot(data, reduction = "umap", group.by = "MenstrualCyclePhase", raster = FALSE)
Fig2b2
ggsave(filename = paste0(dir_out, 'Fig2b2.pdf'),
       plot = Fig2b2,
       width =20, height=9)

```

Supplementary Data Figure 1c
```{r}
SFig1c <- FeaturePlot(data, features = "CycleDay", raster = FALSE)
SFig1c
ggsave(filename = paste0(dir_out, 'SFig1c.pdf'),
       plot = SFig1c,
       width =20, height=9)
```

Supplementary Data Figure 1d
```{r}
SFig1d <- FeaturePlot(data, features = "ProgesteroneSerum", raster = FALSE)
SFig1d
ggsave(filename = paste0(dir_out, 'SFig1d.pdf'),
       plot = SFig1d,
       width =20, height=9)

```

Supplementary Data Figure 3e
/home/common/data/output/projects/ENDO/E044/A066/phase_sc_Fig3_Wang_epi_pseudot/evaluation/ #S3c/d
```{r plott umaps of epithelial cells}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A062/E044A050_seed621_filteredWO_ciliatedMesothelialMUC5B.rds")
A003A033 <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")
table(A003A033$AnnotationRefined)
data <- AddMetaData(data,A003A033)
data <- AddMetaData(data, meta_ptime)
data$epithelial.pseudotime <- as.integer(data$epithelial.pseudotime)
data$Fig3_Wang_epi_pseudot[is.na(data$Fig3_Wang_epi_pseudot)] <- 1 #checked samples

#General Plotting
p1 <- FeaturePlot(data, features = "ProgesteroneSerum", split.by = "EndometriosisStatus")
p2 <- DimPlot(data, group.by = "EndometriosisStatus", split.by = "CyclePhase")
p3 <- DimPlot(data, split.by = "Clusters", group.by = "EndometriosisStatus")
#plot_grid(p1,p2,p3, ncol=1)
ggsave(filename = paste0(dir_out, 'EpiUMAP_general.pdf'),
       plot = plot_grid(p1,p2,p3, ncol=1),
       width =8 ,height=10)

#Plotting of mens cycle info
p1 <- DimPlot(data, group.by = "Fig3_Wang_epi_pseudot")
p2 <- DimPlot(data, group.by = "MenstrualCyclePhase")

#FeaturePlot(seurat_obj, features = "trajectory_day567_epithelial.pseudotime")
#plot_grid(p1,p2,p3,p4,p5,p6, ncol=3)

ggsave(filename = paste0(dir_out, 'EpiUMAP_PhasePrediction_w11.pdf'),
       plot = plot_grid(p1,p2, ncol=2),
       width =11 ,height=5)

```

```{r plotting umaps of entire dataset}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")
A003A033 <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")
table(A003A033$AnnotationRefined)
data <- AddMetaData(data,A003A033)
E031A004meta <- read.csv("/home/common/data/output/projects/ENDO/E044/A033/E031A004_proliferative_SCT_percent.mt_sample_integration_batch_hvg_annotated_meta.csv")
E031A004meta$cell_id <- paste(E031A004meta$sample, E031A004meta$cell.names, sep = "_")
rownames(E031A004meta) <- E031A004meta$cell_id 
data <- AddMetaData(data, E031A004meta[, "Annotation", drop = FALSE])

#Plot clusters, batch,sample EndometriosisStatus, cycle phase
p1 <- DimPlot(data, reduction = "umap", group.by = "EndometriosisStatus", raster = FALSE)
p2 <- DimPlot(data, reduction = "umap", group.by = "batch", raster = FALSE)
p3 <- DimPlot(data, reduction = "umap", group.by = "sample", raster = FALSE) +
  theme(legend.text = element_text(size=4))+
  theme(legend.key.size = unit(0.2, 'cm'))+  guides(color = guide_legend(ncol=1) )
p4 <- DimPlot(data, reduction = "umap", group.by = "celline", raster = FALSE)
p5 <- DimPlot(data, reduction = "umap", group.by = "MenstrualCyclePhase", raster = FALSE)
p6 <- DimPlot(data, reduction = "umap", group.by = "AnnotationMain", raster = FALSE)
p7 <- DimPlot(data, reduction = "umap", group.by = "Symphony_Refined_A029", raster = FALSE)
p8 <- DimPlot(data, reduction = "umap", group.by = "CyclePhase", raster = FALSE)
p9 <- FeaturePlot(data, features = "ProgesteroneSerum", raster = FALSE)
ggsave(filename = paste0(dir_out, 'UMAPs_p1p9.pdf'),
       plot = plot_grid(p1,p2,p3,p4,p6,p7,p5,p8,p9, ncol=3),
       width =28 ,height=24)
ggsave(filename = paste0(dir_out, 'UMAPs_p1p9.png'),
       plot = plot_grid(p1,p2,p3,p4,p6,p7,p5,p8,p9, ncol=3),
       width =28 ,height=24)

p10 <- DimPlot(data, reduction = "umap", group.by = "AnnotationMain", raster = FALSE)
p11 <- FeaturePlot(data, reduction = "umap", features = "Symphony_Global_prob", raster = FALSE)
p12 <- DimPlot(data, reduction = "umap", group.by = "Symphony_Refined", raster = FALSE)
p13 <- FeaturePlot(data, reduction = "umap", features = "Symphony_Refined_prob", raster = FALSE)
p14 <- DimPlot(data, reduction = "umap", group.by = "AnnotationRefined", raster = FALSE)
ggsave(filename = paste0(dir_out, 'UMAPs_p10p14.pdf'),
       plot = plot_grid(p10,p11,p12,p13,p14, ncol=3),
       width =28 ,height=16)

p20 <- DimPlot(data, reduction = "umap", group.by = "batch", raster = FALSE)
p21 <- FeaturePlot(data, reduction = "umap", features = "ProgesteroneSerum", raster = FALSE)
p22 <- FeaturePlot(data, reduction = "umap", features = "CycleDay", raster = FALSE)
p23 <- DimPlot(data, reduction = "umap", group.by = "MenstrualCyclePhase", raster = FALSE)

#Order data by main5
###In progress
order <- A003A033 %>%
  select(AnnotationMain, AnnotationRefined) %>%
  filter(!is.na(AnnotationMain)) %>%
  distinct() %>%
  arrange(AnnotationRefined) %>%
  mutate(AnnotationMain = factor(AnnotationMain, levels = c( "endothelial","stromal","lymphocyte","epithelial","myeloid"))) %>%
  arrange(AnnotationMain)
data$AnnotationRefined <- factor(x = data$AnnotationRefined, levels = order$AnnotationRefined)
p24 <- DimPlot(data, reduction = "umap", group.by = "AnnotationRefined",  raster = FALSE)
p24
data$Fig3_Wang_epi_pseudot[is.na(data$Fig3_Wang_epi_pseudot)] <- 1 #checked samples
p25 <- DimPlot(data, reduction = "umap", group.by = "Fig3_Wang_epi_pseudot",na.value = "",  raster = FALSE)
p25
#Highlight cycling fibroblasts
Idents(data) <- data$Annotation
table(data$Annotation)
cyclingfibs <- WhichCells(data, idents = c("Cycling Fibroblasts"))
p26 <- DimPlot(data, reduction = "umap", cells.highlight = cyclingfibs, cols.highlight = c("darkred"), cols= "grey", raster = FALSE)
p26

plot_grid(p20,p21,p22,p23, ncol=2)
ggsave(filename = paste0(dir_out, 'UMAPs_p20p23.pdf'),
       plot = plot_grid(p20,p21,p22,p23, ncol=2),
       width =24, height=18)
ggsave(filename = paste0(dir_out, 'UMAPs_p20p23.png'),
       plot = plot_grid(p20,p21,p22,p23, ncol=2),
       width =24, height=18)
ggsave(filename = paste0(dir_out, 'UMAPs_p24.pdf'),
       plot = p24,
       width =20, height=9)
ggsave(filename = paste0(dir_out, 'UMAPs_p24.png'),
       plot = p24,
       width =20, height=9)
ggsave(filename = paste0(dir_out, 'UMAPs_p25.pdf'),
       plot = p25,
       width =13, height=9)
ggsave(filename = paste0(dir_out, 'UMAPs_p25.png'),
       plot = p25,
       width =13, height=9)
ggsave(filename = paste0(dir_out, 'UMAPs_p26.pdf'),
       plot = p26,
       width =13, height=9)
ggsave(filename = paste0(dir_out, 'UMAPs_p26.png'),
       plot = p26,
       width =13, height=9)

#Plot objects with split.by sample
p15 <- DimPlot(data, split.by = "sample", raster = FALSE)
ggsave(plot = p15, filename = paste0(dir_out,'A003_samples.pdf'),width =45 ,height=5)

#Plot objects with split.by EndometriosisStatus
p16 <- DimPlot(data, split.by = "EndometriosisStatus", raster = FALSE)
ggsave(plot = p16, filename = paste0(dir_out,'A003_EndometriosisStatus.pdf'),width =15 ,height=5)

rm(data)
gc()
```

Supplementary Data Figure 3 g
```{r}
####### Download data from Tan et al. 2022 (DOI: 10.1038/s41556-022-00961-5) through the command line
#wget --no-check-certificate https://singlecell.jax.org/datasets/endometriosis/h5ad/endo-2022_epithelial.h5ad

####### Convert .h5ad files to .h5seurat files
library(SeuratDisk)
Convert("endo-2022_epithelial.h5ad", dest = "h5seurat", assay = "RNA", overwrite = TRUE)

Tan_epithelial <- LoadH5Seurat(paste0("endo-2022_epithelial.h5seurat"), assays = "RNA")

Tan_epithelial <- ScaleData(object = Tan_epithelial, features = data@assays[["RNA"]]@data@Dimnames[[1]])

#Subset cells from the endometrium (exclude cells from endometriosis lesions)
Idents(Tan_epithelial) <- Tan_epithelial$sample_type_rename
Tan_epithelial <- subset(Tan_epithelial, idents = c("Ctrl", "EuE"))

# factor levels in better order (E10 and E11 don't have endometrium cells)
Tan_epithelial$sample <- factor(Tan_epithelial$sample, levels = c("C01", "C02", "C03", "E01", "E02", "E03", "E04", "E05", "E06", "E07", "E08", "E09"))

# Generate heatmap of the endometrium cells from Tan et al. 2022 with the same epithelial menstrual marker genes as in our manuscript Figure 2d, for comparison
feature_list_figure2 <- c("SFRP4", "LAMC2", "MMP11", "MMP7", "SPRY1", "TFPI2","STEAP4","ADAMTS8","CAPN6","SLC37A2","SCGB1D2", "MT1F", "MT1X","RIMKLB", "S100P", "DEFB1","G0S2","NNMT","GPX3", "PAEP")

heatmap <- DoHeatmap(
  object = Tan_epithelial,
  features = intersect(feature_list_figure2,names),
  cells = NULL,
  group.by = "sample",
  assay = "RNA",
  group.bar = TRUE) + 
  scale_fill_gradientn(limits = c(-2, 2), colors = c("#FF00FF", "#000000", "#FFFF00"))
heatmap
ggsave(
    filename = paste0(dir_out, "Tan_epithelial_Fig2EpiMarkers.pdf"),
    plot = last_plot(),
    width = 12.3,
    height = 5)

#As control also plot menstrual marker genes from Wang et al. 2020 (DOI: 10.1038/s41591-020-1040-z)
feature_list_Wang <- c("PLAU", "MMP7", "THBS1", "CADM1", "NPAS3", "ATP1A1", "ANK3", "ALPL", "TRAK1", "SCGB1D2", "MT1F", "MT1X", "MT1E", "MT1G", "CXCL14", "MAOA", "DPP4", "NUPR1", "GPX3", "PAEP")

heatmap <- DoHeatmap(
  object = Tan_epithelial,
  features = intersect(feature_list_Wang, names),
  cells = NULL,
  group.by = "sample",
  assay = "RNA",
  group.bar = TRUE) + 
  scale_fill_gradientn(limits = c(-2, 2), colors = c("#FF00FF", "#000000", "#FFFF00"))
heatmap
ggsave(
    filename = paste0(dir_out, "Tan_epithelial_WangEpiMarkers.pdf"),
    plot = last_plot(),
    width = 15,
    height = 15)

#The menstrual marker gene expression distribution of each sample looks comparable to our manuscript figure 2d, confirming that the different samples in Tan et al. 2022 express menstrual marker genes from different menstrual cycle phases

```

Supplementary Data Figure 3 h
```{r}
####### Download data from Tan et al. 2022 (DOI: 10.1038/s41556-022-00961-5) through the command line
#wget --no-check-certificate https://singlecell.jax.org/datasets/endometriosis/h5ad/endo-2022_stromal.h5ad

####### Convert .h5ad files to .h5seurat files
library(SeuratDisk)
Convert("endo-2022_stromal.h5ad", dest = "h5seurat", assay = "RNA", overwrite = TRUE)
Tan_stromal <- LoadH5Seurat(paste0("endo-2022_stromal.h5seurat"), assays = "RNA")

# 
FEATURES <- data@assays[["RNA"]]@data@Dimnames[[1]]
Tan_stromal <- ScaleData(object = Tan_stromal, features = FEATURES)

#Subset cells from the endometrium (exclude cells from endometriosis lesions)
Idents(Tan_stromal) <- Tan_stromal$sample_type_rename
Tan_stromal <- subset(Tan_stromal, idents = c("Ctrl", "EuE"))

# factor levels in better order (E10 and E11 don't have endometrium cells)
Tan_stromal$sample <- factor(Tan_stromal$sample, levels = c("C01", "C02", "C03", "E01", "E02", "E03", "E04", "E05", "E06", "E07", "E08", "E09"))

# Generate heatmap of the endometrium cells from Tan et al. 2022 with the same stromal menstrual marker genes as in our manuscript Figure 2d, for comparison
feature_list_figure2_stromal <- c("SFRP1", "H1-1", "H3C2", "WNT5A","CILP","BRINP1", "CFD","PTGDS","SLIT3","RPRM", "TIMP3", "IGFBP1","LEFTY2","LRRC15","LUM")

#Generate heatmap

heatmap <- DoHeatmap(
  object = Tan_stromal,
  features = intersect(feature_list_figure2_stromal,FEATURES),
  cells = NULL,
  group.by = "sample",
  assay = "RNA",
  group.bar = TRUE) + 
  scale_fill_gradientn(limits = c(-2, 2), colors = c("#FF00FF", "#000000", "#FFFF00"))
heatmap
ggsave(
    filename = paste0(dir_out, "Tan_stromal_Fig2stromalMarkers.pdf"),
    plot = last_plot(),
    width = 12.3,
    height = 5)

#Wang genes
feature_list_Wang_stromal <- c('STC1', 'NFATC2', 'BMP2' ,'PMAIP1' ,'MMP11' ,'SFRP1', 'WNT5A' ,'ZFYVE21' ,'CILP' ,'SLF2' ,'MATN2', 'S100A4' ,'DKK1' ,'CRYAB', 'FOXO1', 'IL15', 'FGF7', 'LMCD1')

heatmap <- DoHeatmap(
  object = Tan_stromal,
  features = intersect(feature_list_Wang_stromal, FEATURES),
  cells = NULL,
  group.by = "sample",
  assay = "RNA",
  group.bar = TRUE) + 
  scale_fill_gradientn(limits = c(-2, 2), colors = c("#FF00FF", "#000000", "#FFFF00"))
heatmap
ggsave(
    filename = paste0(dir_out, "Tan_stromal_WangstromalMarkers.pdf"),
    plot = last_plot(),
    width = 15,
    height = 15)

#The menstrual marker gene expression distribution of each sample looks comparable to our manuscript figure 2d, confirming that the different samples in Tan et al. 2022 express menstrual marker genes from different menstrual cycle phases

```

Supplementary Data Figure 1a and 1b
```{r VlnPlots of nCounts and nFeatures}
##VlnPlots for "nFeature_RNA", "nCount_RNA","nCount_SCT","nFeature_SCT","percent.mt"
VlnPlot(data,
  features = c("nFeature_RNA", "nCount_RNA","nCount_SCT","nFeature_SCT","percent.mt"),
  group.by='EndometriosisStatus', ncol = 1, pt.size=0)
ggsave(paste0(dir_out, "VlnPlot_GroupByEndometriosisStatus_nFeature_nCounts_RNA_SCT_percMT.pdf"),
  plot = last_plot(),
  width = 4, height = 13)

##VlnPlots grouped by sample of "nFeature_RNA", "nCount_RNA","nCount_SCT","nFeature_SCT","percent.mt" 
# default level sort order is alphabetical, therefore refactor before plotting
data[["EndometriosisStatus"]] <- factor(unlist(data[["EndometriosisStatus"]])) 
#VlnPlot
VlnPlot(data,
  features = c("nFeature_RNA", "nCount_RNA","nCount_SCT","nFeature_SCT","percent.mt"),
  group.by='sample', ncol = 1, pt.size=0)
ggsave(paste0(dir_out, "VlnPlot_GroupBySample_nFeature_nCounts_RNA_SCT_percMT.pdf"),
  plot = last_plot(),
  width = 25, height = 15)
#Plot VlnPlots with boxplots and save them
VlnPlot(data, features = "nCount_RNA", pt.size = 0) + geom_boxplot(width=0.1,fill="white",outlier.size=0)
VlnPlot(data, features = "nFeature_RNA", pt.size = 0) + geom_boxplot(width=0.1,fill="white",outlier.size=0)
VlnPlot(data, features = "nCount_SCT", pt.size = 0) + geom_boxplot(width=0.1,fill="white",outlier.size=0)
VlnPlot(data, features = "nFeature_SCT", pt.size = 0) + geom_boxplot(width=0.1,fill="white",outlier.size=0)
VlnPlot(data, features = "percent.mt", pt.size = 0) + geom_boxplot(width=0.1,fill="white",outlier.size=0)

```

PCA 
A072 1e, 2a, S1e, S3a
A082 S1f

```{r different counts and means/medians}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_metadata.rds")

#Percentage of main 5
table <- as.data.frame(table(data$AnnotationMain))
table$Percentage <- round((table$Freq / sum(table$Freq)) * 100, 1)

#Number of cells for CTL and Endo
table(data$EndometriosisStatus)

#Number of patients for CTL and Endo
data %>% select(sample,EndometriosisStatus) %>%
  distinct() %>%
  group_by(EndometriosisStatus) %>%
  summarize(number_of_rows = n())

#mean number of cells per sample
mean(table(data$sample))

#median count/feature
median_summary <- data %>%
  select(nCount_RNA, nCount_SCT, nFeature_RNA, nFeature_SCT) %>%
  summarize(
    median_nCount_RNA = median(nCount_RNA),
    median_nCount_SCT = median(nCount_SCT),
    median_nFeature_RNA = median(nFeature_RNA),
    median_nFeature_SCT = median(nFeature_SCT)
  )

```

Supplementary Data Figure 2
```{refined cluster markers}
#Load data
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")
meta <- data@meta.data

#Set SCT as default and remove RNA to save space
DefaultAssay(data) <- "SCT"
data@assays$RNA <- NULL
gc()

#Remove data without annotation (no NAs!)
Idents(data) <- data$Symphony_Refined
data <- subset(data, subset = Symphony_Refined %in% levels(as.factor(data$Symphony_Refined)))
gc()

#setdiff(pops_to_plot,unique(data$Symphony_Refined))
######################################################
#Tan2022 Sup Fig. 7c, epithelial
genes_to_plot <- c("TPPP3", "PIFO", "FOXJ1", "TP63", "KRT5", "SOX9", "LGR5", "MUC5B", "TFF3", "RUNX3", "SAA1", "SIX1", "PROM1", "PROM2", "ANPEP", "KIAA1324", "SPDEF", "SMAD9", "CD36", "HSD17B2", "SCGB2A2", "GDA", "FGFR2", "WNT7A", "FGF9", "PTGS1", "MSLN", "UPK3B", "CALB2", "PGR", "ESR1", "MMP7", "TIMP1", "IDO1", "PSAT1", "ENPP3", "GNG11", "CREB3L1", "IHH", "TOP2A", "MT1F", "MT1E", "MT1X", "SERPINA5", "SPP1", "PAEP", "CXCL14", "DPP4")
pops_to_plot <- c("mesothelial", "ciliated", "mid-secretory", "lumenal 2", "lumenal 1", "lumenal", "glandular early-secretory", "glandular", "TP63+/KRT5+", "MUC5B+")

# reorder clusters
pops_seurat <- subset(data, idents = pops_to_plot)
pops_seurat@active.ident <- factor(pops_seurat@active.ident, 
                            levels=pops_to_plot)
p <- Seurat::DotPlot(
  object = pops_seurat,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = p, "./dotplots/Tan2022Fig3b_epithelial_SCT.pdf", width = 17, height = 7)
######################################################
#DotPlot of features from Tan2022 Fig. 4b, Lymphocytes
genes_to_plot <- c("FCGR3A","KLRD1","ITGA1","LILRB1","ITGAX","KLRC1","KLRC2","EPAS1","KIR2DL4","IGFBP2","CD160","GZMH","KLRF1","FGFBP2","TBX21","CD8A","CD8B","CRTAM","CCR5","NCR3","RORC","IL7R","TCF7","CD4","CCR7","SELL","LEF1","FOXP3","CTLA4","IL2RA","CXCR6","ZNF683","ITGAE","XCL1","SLC4A10","DPP4","GZMK","IGKC", "MZB1", "DERL3", "CD79A", "CD79B", "MS4A1", "BANK1")
pops_to_plot <- c(meta %>% filter(AnnotationMain == "lymphocyte") %>% pull(Symphony_Refined) %>% unique())

# reorder clusters
pops_seurat <- subset(data, idents = pops_to_plot)
pops_seurat@active.ident <- factor(pops_seurat@active.ident, 
                           levels=c("B cell", "plasma", "CD8 T$_{RM}$","CD4 T$_{RM}$","T$_{Reg}$", "T$_N$/T$_{CM}$" , "ILC" , "T$_{EM}$","CTL", "pNK", "NK3","NK2","NK1"))

p <- Seurat::DotPlot(
  object = pops_seurat,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = p, "./dotplots/Tan2022Fig3b_Lymphocytes_SCT.pdf", width = 17, height = 7)
######################################################
#Tan2022 Fig. 3b, stromal
genes_to_plot <- c("RERGL", "MCAM", "MYH11", "STEAP4","GGT5", "RGS5")
pops_to_plot <- c("Prv-MYH11", "Prv-CCL19", "Prv-STEAP4", "VSMC")
# reorder clusters
pops_seurat <- subset(data, idents = pops_to_plot)
pops_seurat@active.ident <- factor(pops_seurat@active.ident, 
                            levels=c("VSMC", "Prv-STEAP4", "Prv-CCL19","Prv-MYH11"))

p <- Seurat::DotPlot(
  object = pops_seurat,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = p, "./dotplots/Tan2022Fig3b_stromal_SCT.pdf", width = 7, height = 7)
######################################################
#Tan2022 Fig. 4b, macrophages
genes_to_plot <- c("FLOR2", "LYVE1", "MRC1", "CX3CR1", "ICAM2", "CLEC5A", "CCR2", "VEGFA", "FN1", "SPARC", "APOE", "SPP1")
pops_to_plot <- c(meta %>% filter(AnnotationMain == "myeloid") %>% pull(Symphony_Refined) %>% unique())
pops_to_plot <- c("M$\\Phi$5-activated","M$\\Phi$4-infiltrated","M$\\Phi$3-APOE","M$\\Phi$2-peritoneal","M$\\Phi$1-LYVE1")

# reorder clusters
pops_seurat <- subset(data, idents = pops_to_plot)
pops_seurat@active.ident <- factor(pops_seurat@active.ident, 
                            levels=pops_to_plot)
p <- Seurat::DotPlot(
  object = pops_seurat,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = p, "./dotplots/Tan2022Fig3b_macrophages_SCT.pdf", width = 10, height = 7)
######################################################
#Tan2022 Fig. 5a, DCs
genes_to_plot <- c("CD1C", "CLEC9A", "LAMP3")
pops_to_plot <- c("mDC", "cDC1", "DC3", "cDC2", "pre-cDC2")
# reorder clusters
pops_seurat <- subset(data, idents = pops_to_plot)
pops_seurat@active.ident <- factor(pops_seurat@active.ident, 
                            levels=pops_to_plot)
p <- Seurat::DotPlot(
  object = pops_seurat,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = p, "./dotplots/Tan2022Fig3b_DCs_SCT.pdf", width = 5, height = 7)

rm(data)
gc()

```

```{major cluster markers}
#Load data
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")

#Set SCT as default and remove RNA to save space
DefaultAssay(data) <- "SCT"
data@assays$RNA <- NULL
gc()

#Remove data without annotation
Idents(data) <- data$AnnotationMain
data <- subset(data, subset = AnnotationMain %in% levels(as.factor(data$AnnotationMain)))

GENES <- c("EPCAM","CDH1","PDGFRA","VWF","PECAM1","PTPRC","CD68","CD14","CD2")
p <- VlnPlot(data, features = GENES, group.by = "AnnotationMain",raster=FALSE, pt.size = 0, assay = "SCT", stack=T, flip=T)
ggsave(plot = p, "./VlnPlotMain5_SCT_notds.pdf", width = 6, height = 4.5)

GENES <- c("PDGFRA","DCN","COL1A1","COL6A1","CRISPLD2","COL6A3","LUM","COL5A1") #stomal genes, only "PDGFRA" and "CRISPLD2" are cleanly expressed in stromal only, the others are a bit mixed
p <- VlnPlot(data, features = GENES, group.by = "AnnotationMain",raster=FALSE, pt.size = 0, assay = "RNA", stack=T, flip=T)
ggsave(plot = p, "./VlnPlotMain5_stromalMG_RNA_notds.pdf", width = 6, height = 4.5)

p <- VlnPlot(data, features = GENES, group.by = "AnnotationMain",raster=FALSE, pt.size = 0, assay = "SCT", stack=T, flip=T)
ggsave(plot = p, "./VlnPlotMain5_stromalMG_SCT_notds.pdf", width = 6, height = 4.5)

GENES <- c("EPCAM","CDH1","PDGFRA","COL1A1", "VWF","PECAM1","PTPRC","CD68","CD14","CD2")
p <- VlnPlot(data, features = GENES, group.by = "AnnotationMain",raster=FALSE, pt.size = 0, assay = "RNA", stack=T, flip=T)
ggsave(plot = p, "./VlnPlotMain5_RNA_notds.pdf", width = 6, height = 4.5)

table(data$AnnotationMain)

#Make DotPlots
DefaultAssay(data) <- "SCT"

GENES

######################################################
#Tan2022 main markers
genes_to_plot <- c("EPCAM","CDH1","PDGFRA","COL1A1", "VWF","PECAM1","PTPRC","CD68","CD14","CD2")
pops_to_plot <- c("stromal","epithelial","myeloid","lymphocyte","endothelial")

Seurat::DotPlot(
  object = data,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = last_plot(), "./dotplots/main5_main5markers_SCT.pdf", width = 10, height = 5)

#stomal markers
genes_to_plot <- c("PDGFRA","DCN","COL1A1","COL6A1","CRISPLD2","COL6A3","LUM","COL5A1") #stomal genes, only "PDGFRA" and "CRISPLD2" are cleanly expressed in stromal only, the others are a bit mixed
Seurat::DotPlot(
  object = data,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = last_plot(), "./dotplots/main5_stromalmarkers_SCT.pdf", width = 10, height = 5)

#all markers + stomal
genes_to_plot <- c("EPCAM","CDH1","PDGFRA","COL5A1", "VWF","PECAM1","PTPRC","CD68","CD14","CD2","DCN","COL6A1","CRISPLD2","COL6A3","LUM","COL1A1") #stomal genes, only "PDGFRA" and "CRISPLD2" are cleanly expressed in stromal only, the others are a bit mixed
Seurat::DotPlot(
  object = data,
  scale = T,
  features = genes_to_plot,
  idents = pops_to_plot,
) + theme(axis.text.x = element_text(angle = 90))
ggsave(plot = last_plot(), "./dotplots/main5_allmarkers_SCT.pdf", width = 17, height = 5)

```

```{r barplot distribution of cv splits, all cells, prolif}
meta <- read.csv("/home/common/data/output/projects/ENDO/E042/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv")

library(reshape2)
library(ggplot2)
library(forcats)

# reshape df
melted_meta <- melt(meta, id.vars = c("lane", "endo_EndometriosisGrade"), measure.vars = c("fold1", "fold2", "fold3", "fold4", "fold5"))
## create the bar plot for endo_EndometriosisGrade
ggplot(melted_meta, aes(x = variable, fill = endo_EndometriosisGrade)) +
  geom_bar(position = "stack", width = 0.7) +
  facet_wrap(~ fct_rev(value), ncol = 2) +
  labs(title = "Bar Plot of Samples by Endo EndometriosisGrade",
       x = "Fold",
       y = "Number of Samples") +
  scale_fill_manual(values = c("severe" = "red", "mild" = "orange", "CTL" = "blue")) +
  theme_minimal() +
  coord_flip()
ggsave(paste0(dir_out,"Bar Plot of Samples by Endo EndometriosisGrade_allcells.pdf"), plot = last_plot(), height = 6.3, width = 7.2)

## create the bar plot for lane
ggplot(melted_meta, aes(x = variable, fill = lane)) +
  geom_bar(position = "stack", width = 0.7) +
  facet_wrap(~ fct_rev(value), ncol = 2) +
  labs(title = "Bar Plot of Samples by lane",
       x = "Fold",
       y = "Number of Samples") +
  #scale_fill_manual(values = c("severe" = "red", "mild" = "orange", "CTL" = "blue")) +
  theme_minimal() +
  coord_flip()
ggsave(paste0(dir_out,"Bar Plot of Samples by lane_allcells.pdf"), plot = last_plot(), height = 6.3, width = 7.2)

##Upsetplot of eval samples
df <- meta %>%
  rownames_to_column(var = "name") %>%
  select(-name) %>% 
  select(fold1, fold2, fold3, fold4, fold5) %>%
  mutate_all(~ ifelse(. == 'eval', TRUE, FALSE)) %>%
  filter(rowSums(. == TRUE) > 0)
pdf(file.path(dir_out, "upset_CellCnnsig_all_evalsamples.pdf"))
upset(df,
  colnames(df)) +
  ggtitle("all cells")
dev.off()

################
#heatmap, not working yet

# Extracting the necessary columns from the 'meta' dataframe
data_matrix <- meta[, c("fold1", "fold2", "fold3", "fold4", "fold5")]
data_matrix_numeric <- apply(data_matrix, 2, function(x) as.numeric(x == "eval"))

# Adding annotation columns
annotation_data <- meta[, c("EndometriosisStatus", "lane", "batch", "endo_EndometriosisGrade")]

library(ComplexHeatmap)
library(circlize)

# Define a custom color palette
my_color_palette <- colorRamp2(c(0, 0.5, 1), c("blue", "white", "red"))
png(file=paste0(dir_out,'heatmap.pdf'))
# Create the heatmap using ComplexHeatmap
heatmap <- Heatmap(
  data_matrix_numeric,
  col = my_color_palette,
  row_dend_reorder = TRUE,  # Adjust this based on your preference
  column_title = "Folds",
  name = "Fold Value",
  show_row_names = FALSE,
  show_column_names = TRUE,
  cluster_columns = FALSE
)

draw(heatmap)
dev.off()
```

```{r barplot distribution of cv splits, cf, prolif}
meta <- read.csv("/home/common/data/output/projects/ENDO/E041/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv")

library(reshape2)
library(ggplot2)
library(forcats)

# reshape df
melted_meta <- melt(meta, id.vars = c("lane", "endo_EndometriosisGrade"), measure.vars = c("fold1", "fold2", "fold3", "fold4", "fold5"))
# create the bar plot for endo_EndometriosisGrade
ggplot(melted_meta, aes(x = variable, fill = endo_EndometriosisGrade)) +
  geom_bar(position = "stack", width = 0.7) +
  facet_wrap(~ fct_rev(value), ncol = 2) +
  labs(title = "Bar Plot of Samples by Endo EndometriosisGrade",
       x = "Fold",
       y = "Number of Samples") +
  scale_fill_manual(values = c("severe" = "red", "mild" = "orange", "CTL" = "blue")) +
  theme_minimal() +
  coord_flip()
ggsave(paste0(dir_out,"Bar Plot of Samples by Endo EndometriosisGrade_cf.pdf"), plot = last_plot(), height = 6.3, width = 7.2)

# create the bar plot for lane
ggplot(melted_meta, aes(x = variable, fill = lane)) +
  geom_bar(position = "stack", width = 0.7) +
  facet_wrap(~ fct_rev(value), ncol = 2) +
  labs(title = "Bar Plot of Samples by lane",
       x = "Fold",
       y = "Number of Samples") +
  #scale_fill_manual(values = c("severe" = "red", "mild" = "orange", "CTL" = "blue")) +
  theme_minimal() +
  coord_flip()
ggsave(paste0(dir_out,"Bar Plot of Samples by lane_cf.pdf"), plot = last_plot(), height = 6.3, width = 7.2)

##Upsetplot of eval samples
df <- meta %>%
  rownames_to_column(var = "name") %>%
  select(-name) %>% 
  select(fold1, fold2, fold3, fold4, fold5) %>%
  mutate_all(~ ifelse(. == 'eval', TRUE, FALSE)) %>%
  filter(rowSums(. == TRUE) > 0)
pdf(file.path(dir_out, "upset_CellCnnsig_cf_evalsamples.pdf"))
upset(df,
  colnames(df)) +
  ggtitle("all cells")
dev.off()

################
#heatmap, not working yet

# Extracting the necessary columns from the 'meta' dataframe
data_matrix <- meta[, c("fold1", "fold2", "fold3", "fold4", "fold5")]
data_matrix_numeric <- apply(data_matrix, 2, function(x) as.numeric(x == "eval"))

# Adding annotation columns
annotation_data <- meta[, c("EndometriosisStatus", "lane", "batch", "endo_EndometriosisGrade")]

library(ComplexHeatmap)
library(circlize)

# Define a custom color palette
my_color_palette <- colorRamp2(c(0, 0.5, 1), c("blue", "white", "red"))
png(file=paste0(dir_out,'heatmap.pdf'))
# Create the heatmap using ComplexHeatmap
heatmap <- Heatmap(
  data_matrix_numeric,
  col = my_color_palette,
  row_dend_reorder = TRUE,  # Adjust this based on your preference
  column_title = "Folds",
  name = "Fold Value",
  show_row_names = FALSE,
  show_column_names = TRUE,
  cluster_columns = FALSE
)

draw(heatmap)
dev.off()
```

```{r DimPlot with cycling fibroblasts}
data <- readRDS("/home/common/data/output/projects/ENDO/E031/A004/proliferative_SCT_percent.mt_sample_integration_batch_hvg_annotated.RDS")
DimPlot(data, raster = FALSE)
meta <- data@meta.data
write.csv(meta, paste0(dir_out,"E031A004_proliferative_SCT_percent.mt_sample_integration_batch_hvg_annotated_meta.csv"))

# make dotplot
features <- c("NFKBIA", "SLPI", "CXCL2","CXCL3")
Idents(data) <- data$Annotation

epithelial <- subset(data, idents = "Unciliated")

DotPlot(
  epithelial,
  assay = "SCT",
  features,
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = "EndometriosisStatus",
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
)

```

```{r Expression distribution of all genes and pca signature}
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")

#CellCnn gene signature
#Summary table for cyc fibro corrgpca: 
CellCnnsig_fibro_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E041/A010/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv") 
#Summary table for  cyc fibro pca :  
CellCnnsig_fibro_pca <- read.csv("/home/common/data/output/projects/ENDO/E041/A012/sigs/weighted_score_sig_pca_1pc.csv") 
#Summary table for  all cells corrgpca: 
CellCnnsig_all_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E042/A011/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv") 
#Summary table for  all cells pca: 
CellCnnsig_all_pca <- read.csv("/home/common/data/output/projects/ENDO/E042/A012/sigs/weighted_score_sig_pca_1pc.csv") 
CellCnnsig <- unique(c(CellCnnsig_all_corrgpca$gene, CellCnnsig_all_pca$gene))
ScaiNETsamples <- read.csv("/home/common/data/output/projects/ENDO/E041/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv")

#Make UpSet plot
# Create table with the gene sets
gene_sets <- list(
    PCA = CellCnnsig_all_pca$gene,
    CorrgPCA = CellCnnsig_all_corrgpca$gene)
all_genes <- unique(unlist(gene_sets))
df <- as.data.frame(lapply(gene_sets, function(gene_set) {
  setNames(as.logical(all_genes %in% gene_set), all_genes)
}))
# Make upset plot
pdf(file.path(dir_out, "upset_CellCnnsig_all_pca_corrgpcs.pdf"))
upset(df,
  colnames(df),
  set_sizes=(
        upset_set_size(position='right'))) +
  ggtitle("all cells")
dev.off()

#CTL and Endo samples used for the DEG
#Filter for samples with >50 cells/sample
data_metadata <- data@meta.data
df_sample_cells <- data@meta.data %>% 
  dplyr::group_by(sample, EndometriosisStatus) %>% 
  dplyr::summarize(cells = n()) %>% 
  dplyr::filter(cells >= 50) %>% 
  dplyr::filter(sample %in% ScaiNETsamples$names) 
ENDO <- df_sample_cells %>% filter(EndometriosisStatus == "Endo") %>%
  select(sample) %>%
  pull(sample)
CTL <- df_sample_cells %>% filter(EndometriosisStatus == "CTL") %>%
  select(sample) %>%
  pull(sample)
#Retrieve number of Endo/CTL samples with RNA expression
exprmatrix_data <- AggregateExpression(
  data,
  assays = "RNA",
  features = NULL,
  return.seurat = FALSE,
  group.by = c("sample"),
  slot = "counts",
  verbose = TRUE)[["RNA"]]
exprmatrix_data <- as.data.frame(exprmatrix_data)
exprmatrix_data$gene <- row.names(exprmatrix_data)
exprmatrix_data <- exprmatrix_data %>%
  mutate(Endo_expressed_n = rowSums(select(., all_of(ENDO)) > 0)) %>%
  mutate(CTL_expressed_n = rowSums(select(., all_of(CTL)) > 0)) %>%
  mutate(samples_expressed_n = Endo_expressed_n + CTL_expressed_n) %>%
  mutate(ENDOorCTL_max_expressed_n = pmax(Endo_expressed_n, CTL_expressed_n))
table(exprmatrix_data$ENDOorCTL_max_expressed_n)

exprmatrix_CellCnnsig <- exprmatrix_data %>%
  filter(row.names(exprmatrix_data) %in% CellCnnsig$gene)
table(exprmatrix_CellCnnsig$ENDOorCTL_max_expressed_n)

exprmatrix_data <- exprmatrix_data %>%
  mutate(CellCnnsig = row.names(exprmatrix_data) %in% CellCnnsig$gene)

# Create a histogram
ggplot(exprmatrix_data, aes(x = samples_expressed_n, y = after_stat(count))) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  facet_wrap(~ CellCnnsig, ncol = 2, scales = "free_y") +
  labs(x = "number of samples with gene expression", y = "number of genes expressed") +
  theme_minimal() +
  geom_text(
    aes(label = after_stat(count)),
    stat = "count",
    vjust = -0.5,
    size = 2,
    position = position_stack()
  )

################################ in work
# Assuming your data is stored in a data frame named 'CellCnnsig_all_pca'
library(ggplot2)

color_scale <- scale_fill_gradient(low = "red", high = "green")

heatmap_plot <- ggplot(CellCnnsig_all_pca, aes(x = X, y = gene, fill = rank)) +
  color_scale +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

heatmap_plot <- heatmap_plot +
  geom_text(aes(label = gene), vjust = 1, hjust = 1, size = 3)

# Show the plot
print(heatmap_plot)
#####################################
```

```{r Expression distribution of all genes and pca signature for cf}
#CellCnn gene signature
#Summary table for cyc fibro corrgpca: 
CellCnnsig_fibro_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E041/A010/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv") 
#Summary table for  cyc fibro pca :  
CellCnnsig_fibro_pca <- read.csv("/home/common/data/output/projects/ENDO/E041/A012/sigs/weighted_score_sig_pca_1pc.csv") 
CellCnnsig <- unique(c(CellCnnsig_fibro_corrgpca$gene, CellCnnsig_fibro_pca$gene))

#Make UpSet plot
# Create table with the gene sets
gene_sets <- list(
    PCA = CellCnnsig_fibro_pca$gene,
    CorrgPCA = CellCnnsig_fibro_corrgpca$gene)
all_genes <- unique(unlist(gene_sets))
df <- as.data.frame(lapply(gene_sets, function(gene_set) {
  setNames(as.logical(all_genes %in% gene_set), all_genes)
}))
# Make upset plot
pdf(file.path(dir_out, "upset_CellCnnsig_fibro_pca_corrgpcs.pdf"))
upset(df,
  colnames(df),
  set_sizes=(
        upset_set_size(position='right'))) +
  ggtitle("fibro")
dev.off()
```

```{r Plot DEGs from Tan proliferative, vulcano and DotPlot}
#RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv") %>%
  filter(annotation_reference == "Tan") %>%
  filter(AveExpr > 18.53202)
##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes, there are no hb genes in the DEG list
mt.genes <- grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#mitochondrial genes, no genes
rb.genes <- grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE)#ribosomal genes, 165 genes
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  dplyr::filter(!gene %in% rb.genes) #Focus on protein coding genes
##Rename cluster_id to major cell types
meta <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_metadata.rds")
meta$Symphony_Refined[meta$Symphony_Refined == "M$\\Phi$1-LYVE1"] <- "M-Phi-1-LYVE1"
meta$Symphony_Refined[meta$Symphony_Refined == "glandular early-secretory"] <- "glandular-early-secretory"
meta <- meta %>% 
  select(AnnotationMain,Symphony_Refined,Symphony_Refined_A029_United) %>%
  distinct()
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  mutate(AnnotationMain = cluster_id)
RNAlistRDS005_df$AnnotationMain <- case_when(
  RNAlistRDS005_df$AnnotationMain %in% meta$Symphony_Refined ~ meta$AnnotationMain[match(RNAlistRDS005_df$AnnotationMain, meta$Symphony_Refined)],
  RNAlistRDS005_df$AnnotationMain %in% meta$Symphony_Refined_A029_United ~ meta$AnnotationMain[match(RNAlistRDS005_df$AnnotationMain, meta$Symphony_Refined_A029_United)],
  TRUE ~ RNAlistRDS005_df$AnnotationMain)
RNAlistRDS005_df$AnnotationMain[RNAlistRDS005_df$AnnotationMain == "myeloid" |
                                      RNAlistRDS005_df$AnnotationMain == "lymphocyte"] <- "immune"

###Plot DEGs
#vulcano plot without gene labels
ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc), color = stage)) +
  geom_point() +
  theme_minimal() +
  #ylim(-2, 17) +
  xlim(-1, 1.2)
ggsave(paste0(dir_out,"VulcanoPlot_DEGs_stage.pdf"), 
       plot = last_plot(), 
       height = 6.3, width = 5)

#vulcano plot with gene labels
GenesToLabel <- RNAlistRDS005_df %>% 
               #distinct(gene, logFC,AveExpr,p_adj.loc, .keep_all = TRUE) %>% 
                dplyr::filter(abs(logFC) >= 0.5)
ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc), color = stage)) +
  geom_point() +
  theme_minimal() +
  #ylim(-2, 17) +
  xlim(-1, 1.2) +
  geom_text_repel(data = GenesToLabel,
                  aes(label = gene),
                  size = 2,
                  max.overlaps = 50,
                  box.padding = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50')
#Barplot unique genes per main cell type
RNAlistRDS005_df %>% 
  select(AnnotationMain,gene,stage,cluster_id) %>%
  ggplot(aes(x = stage, fill = cluster_id)) +
  geom_bar() +
  facet_wrap(~ AnnotationMain, ncol = 4) +
  theme(axis.text.x = element_text(angle=45, vjust=.2, hjust=0.3))

#Barplot unique genes per main cell type
RNAlistRDS005_df %>% 
  select(AnnotationMain, gene, stage, cluster_id) %>%
  ggplot(aes(x = stage, fill = factor(AnnotationMain, levels = c("immune", "endothelial", "epithelial", "stromal")))) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.3))

#Barplot unique genes per main cell type
RNAlistRDS005_df %>% 
  select(AnnotationMain, gene, stage) %>%
  distinct() %>%
  ggplot(aes(x = stage, fill = factor(AnnotationMain, levels = c("immune", "endothelial", "epithelial", "stromal")))) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.3))
ggsave(paste0(dir_out,"BarPlot_DEGs_main4.pdf"), 
       plot = last_plot(), 
       height = 4, width = 10)

RNAlistRDS005_df %>% 
      select(AnnotationMain, gene, stage) %>%
      distinct() %>%
      count(stage, AnnotationMain)

RNAlistRDS005_df %>% 
      filter(logFC >= 0.5) %>%
      distinct(gene, stage) %>%
      count(stage)

RNAlistRDS005_df %>% 
      select(gene, stage) %>%
      distinct() %>%
      count(stage)

table <- RNAlistRDS005_df %>% 
      select(AnnotationMain, gene, stage) %>%
      distinct() %>%
      count(AnnotationMain, stage)

```

```{r DEGs vs CellCNN gene signature}
##Load dataframes
#CellCnn gene signature
#Summary table for cyc fibro corrgpca: 
CellCnnsig_fibro_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E041/A010/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv") 
#Summary table for  cyc fibro pca :  
CellCnnsig_fibro_pca <- read.csv("/home/common/data/output/projects/ENDO/E041/A012/sigs/weighted_score_sig_pca_1pc.csv") 
#Summary table for  all cells corrgpca: 
CellCnnsig_all_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E042/A011/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv") 
CellCnnsig_all_corrgpca$origin <- "A011corrgpca"
#Summary table for  all cells pca: 
CellCnnsig_all_pca <- read.csv("/home/common/data/output/projects/ENDO/E042/A012/sigs/weighted_score_sig_pca_1pc.csv") 
CellCnnsig_all_pca$origin <- "A012pca"

CellCnnsig <- rbind(CellCnnsig_all_corrgpca, CellCnnsig_all_pca)

#RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv")

##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
#Haemoglobin genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes
CellCnnsig[CellCnnsig$gene %in% hb.genes, ]#no hb genes
RNAlistRDS005_df[RNAlistRDS005_df$gene %in% hb.genes, ]#no hb genes
#mitochondrial genes
grep("^MT-", CellCnnsig$gene, value=TRUE)#no mt genes
grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#no mt genes
#ribosomal genes
rb.genes_CellCnnsig <- grep(pattern = "^RP[SL]", CellCnnsig$gene, value=TRUE)#3 rb genes
CellCnnsig <- CellCnnsig[!(CellCnnsig$gene %in% rb.genes_CellCnnsig), ]
rb.genes_DEGs <- unique(grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE))#77 rb genes
RNAlistRDS005_df <- RNAlistRDS005_df[!(RNAlistRDS005_df$gene %in% rb.genes_DEGs), ]

##Combine the two dataframes
RNAlistRDS005_df$CellCnnsig <- factor(RNAlistRDS005_df$gene %in% CellCnnsig$gene,
                                        levels = c(FALSE, TRUE),
                                        labels = c("Non-overlapping", "overlapping"))

##Make plots
# Create a scatter plot logFC vs AveExpr
ggplot(RNAlistRDS005_df, aes(x = AveExpr, y = logFC)) +
  geom_point(data = RNAlistRDS005_df, aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = RNAlistRDS005_df %>% dplyr::filter(CellCnnsig == "overlapping"),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "Scatter Plot",
       color = "Overlap with CellCnnsig") +
  theme_minimal()
ggplot(RNAlistRDS005_df, aes(x = AveExpr, y = logFC, color = cluster_id)) +
  geom_point(size = 1) +
  labs(title = "Scatter Plot") +
  theme_minimal()

# Create a scatter plot logFC vs AveExpr
ggplot(RNAlistRDS005_df, aes(x = p_adj.loc, y = logFC)) +
  geom_point(data = RNAlistRDS005_df, aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = RNAlistRDS005_df %>% dplyr::filter(CellCnnsig == "overlapping"),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "Scatter Plot",
       color = "Overlap with CellCnnsig") +
  theme_minimal()
ggplot(RNAlistRDS005_df, aes(x = p_adj.loc, y = logFC, color = cluster_id)) +
  geom_point(size = 1) +
  labs(title = "Scatter Plot") +
  theme_minimal()

# Create a scatter plot
ggplot(RNAlistRDS005_df, aes(x = logFC, y = p_adj.loc)) +
  geom_point(data = RNAlistRDS005_df, aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = RNAlistRDS005_df %>% dplyr::filter(CellCnnsig == "overlapping"),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "Scatter Plot",
       color = "Overlap with CellCnnsig E042A011A012") +
  theme_minimal()

# Create vulcano plot
ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = RNAlistRDS005_df, aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = RNAlistRDS005_df %>% dplyr::filter(CellCnnsig == "overlapping"),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "Scatter Plot",
       color = "Overlap with CellCnnsig E042A011A012") +
  theme_minimal()
ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc), color = cluster_id)) +
  geom_point(size = 1.2) 
ggsave(filename = paste0(dir_out, 'vulcano_clusterid_10_5.pdf'), width =6,height=3)

overlap <- RNAlistRDS005_df[(RNAlistRDS005_df$gene %in% CellCnnsig$gene),]
length(unique(overlap$gene))
#Plot 
par(mfrow = c(2, 1))
plot(density(overlap$AveExpr))
plot(density(RNAlistRDS005_df$AveExpr))

#Histogram
# Create the barplot
summary_data <- RNAlistRDS005_df %>%
  group_by(cluster_id, stage, CellCnnsig) %>%
  dplyr::summarize(Freq = n())
# Create the barplot with color by CellCnnsig

# Create the barplot with a log2 scale on the y-axis
ggplot(summary_data, aes(x = cluster_id, y = Freq, fill = CellCnnsig)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4) +
  scale_y_break(c(100, 150), scales = 0.3)+ 
  scale_y_break(c(350, 2500), scales = 0.3)

ggplot(RNAlistRDS005_df, aes(x = cluster_id, y = p_adj.loc, color = CellCnnsig)) +
  geom_violin() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4)
ggplot(RNAlistRDS005_df, aes(x = cluster_id, y = logFC, color = CellCnnsig)) +
  geom_violin() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4)
ggplot(RNAlistRDS005_df, aes(x = cluster_id, y = AveExpr, color = CellCnnsig)) +
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=0, notch=FALSE) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4)

ggplot(summary_data, aes(x = cluster_id, y = Freq, fill = CellCnnsig)) +
  geom_bar(position = "fill", stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4)

#Get number of overlapping DEGs per cluster
head(RNAlistRDS005_df)
overlapping <- RNAlistRDS005_df[RNAlistRDS005_df$gene %in% CellCnnsig$gene,]
table(overlapping$cluster_id)
table(RNAlistRDS005_df$cluster_id)

#Density plot for logFC
p1 <- ggplot(RNAlistRDS005_df, aes(x = logFC, fill = CellCnnsig)) +
  geom_density(alpha = 0.5) +  
  theme_minimal() + 
  scale_fill_manual(values = c("Non-overlapping" = "blue", "overlapping" = "orange"),drop = FALSE) 
p1
#Density plot for AveExpr
p2 <- ggplot(RNAlistRDS005_df, aes(x = log2(AveExpr - 17), fill = CellCnnsig)) +
  geom_density(alpha = 0.5) + 
  theme_minimal() + 
  scale_fill_manual(values = c("Non-overlapping" = "blue", "overlapping" = "orange"),drop = FALSE) +
  geom_vline(xintercept = 0.5849625)
p2
#Density plot for p_adj.loc
p3 <- ggplot(RNAlistRDS005_df, aes(x = p_adj.loc, fill = CellCnnsig)) +
  geom_density(alpha = 0.5) + 
  theme_minimal() + 
  scale_fill_manual(values = c("Non-overlapping" = "blue", "overlapping" = "orange"),drop = FALSE) 
plot_grid(p1,p2,p3, ncol=1)

```

```{r Make rank by padj.loc and logFC and evaluate in vulcano plot}
table <- RNAlistRDS005_df %>%
  mutate(logFC_rank = rank(-abs(logFC))) %>%
  mutate(p_adj.loc_rank = rank(p_adj.loc)) %>%
  mutate(padj.loc_logFC_rank = (logFC_rank + p_adj.loc_rank)) %>%
  mutate(padj.loc_logFC10_rank = (10*logFC_rank + p_adj.loc_rank)) %>%
  mutate(padj.loc_logFC100_rank = (100*logFC_rank + p_adj.loc_rank)) %>%
  arrange(AveExpr)
table_overlapping <- table[table$gene %in% CellCnnsig$gene,]

#table200 <- table %>% arrange(mean_rank) %>% slice_head(n = 300)

# Create vulcano plot
p1 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC_rank, slice_head300") +
  theme_minimal()
p1

p2 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC10_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC10_rank, slice_head300") +
  theme_minimal()
p2

p3 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC100_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC100_rank, slice_head300") +
  theme_minimal()
p3

p4 <- ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = RNAlistRDS005_df, aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = RNAlistRDS005_df %>% dplyr::filter(CellCnnsig == "overlapping"),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "Scatter Plot",
       color = "CellCnnsigE042A011A012") +
  theme_minimal()
p4

p5 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.3, 25, by = 5),
                       limits = c(18.3, 25),
                       labels = as.character(seq(18.3, 25, by = 5)))
p5
p6 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.5, 25, by = 5),
                       limits = c(18.5, 25),
                       labels = as.character(seq(18.5, 25, by = 5)))
p6
plot_grid(p1,p2,p3,p4,p5,p6, ncol=3)

#Keep in mind to remove gene duplicates before giving ShSh the list
```

```{r Make rank by padj.loc and logFC and AveExpr and evaluate in vulcano plot}
table <- RNAlistRDS005_df %>%
  mutate(logFC_rank = rank(-abs(logFC))) %>%
  mutate(p_adj.loc_rank = rank(p_adj.loc)) %>%
  mutate(AveExpr_rank = rank(-AveExpr)) %>%
  mutate(padj.loc_logFC_AveExpr_rank = (logFC_rank + p_adj.loc_rank + AveExpr_rank)) %>%
  mutate(padj.loc_logFC10_AveExpr_rank = (10*logFC_rank + p_adj.loc_rank + AveExpr_rank)) %>%
  mutate(padj.loc_10logFC_10AveExpr_rank = (logFC_rank + p_adj.loc_rank + 10*AveExpr_rank)) %>%
  arrange(AveExpr)
table_overlapping <- table[table$gene %in% CellCnnsig$gene,]
#table200 <- table %>% arrange(mean_rank) %>% slice_head(n = 300)

# Create vulcano plot
p1 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC_AveExpr_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC_AveExpr_rank, slice_head300") +
  theme_minimal()
p1

p2 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC10_AveExpr_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC10_AveExpr_rank, slice_head300") +
  theme_minimal()
p2

p3 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_10logFC_10AveExpr_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_10logFC_10AveExpr_rank, slice_head300") +
  theme_minimal()
p3

p4 <- ggplot(table_overlapping, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "table_overlapping") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.3, 25, by = 5),
                       limits = c(18.3, 25),
                       labels = as.character(seq(18.3, 25, by = 5)))
p4

p5 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.3, 25, by = 5),
                       limits = c(18.3, 25),
                       labels = as.character(seq(18.3, 25, by = 5)))
p5
p6 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.5, 25, by = 5),
                       limits = c(18.5, 25),
                       labels = as.character(seq(18.5, 25, by = 5)))
p6
plot_grid(p1,p2,p3,p4,p5,p6, ncol=3)

#Keep in mind to remove gene duplicates before giving ShSh the list
```

```{r Make rank by padj.loc and logFC, AveExpr > 18.5 and evaluate in vulcano plot}
table <- RNAlistRDS005_df %>%
  dplyr::filter(AveExpr > 18.5) %>%
  mutate(logFC_rank = rank(-abs(logFC))) %>%
  mutate(p_adj.loc_rank = rank(p_adj.loc)) %>%
  mutate(padj.loc_logFC_rank = (logFC_rank + p_adj.loc_rank)) %>%
  mutate(padj.loc_logFC10_rank = (10*logFC_rank + p_adj.loc_rank)) %>%
  mutate(padj.loc_logFC2_rank = (2*logFC_rank + p_adj.loc_rank)) %>%
  arrange(AveExpr)
table_overlapping <- table[table$gene %in% CellCnnsig$gene,]

#Histogram
# Create the barplot
summary_data <- table %>%
  group_by(cluster_id, stage, CellCnnsig) %>%
  dplyr::summarize(Freq = n())
# Create the barplot with color by CellCnnsig

# Create the barplot with a log2 scale on the y-axis
ggplot(summary_data, aes(x = cluster_id, y = Freq, fill = CellCnnsig)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~stage, scales = "free_x", ncol = 4) +
  scale_y_break(c(40, 50), scales = 0.3)

#table200 <- table %>% arrange(mean_rank) %>% slice_head(n = 300)

# Create vulcano plot
p1 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC_rank, slice_head300") +
  theme_minimal()
p1

p2 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC10_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC10_rank, slice_head300") +
  theme_minimal()
p2

p3 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = table,
             aes(color = "Non-overlapping"), size = 1) +
  geom_point(data = table %>% arrange(padj.loc_logFC2_rank) %>% slice_head(n = 300),
             aes(color = "overlapping"), size = 2) +
  scale_color_manual(values = c("Non-overlapping" = "black", "overlapping" = "purple")) +
  labs(title = "padj.loc_logFC2_rank, slice_head300") +
  theme_minimal()
p3

p4 <- ggplot(table_overlapping, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "table_overlapping") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.5, 25, by = 5),
                       limits = c(18.5, 25),
                       labels = as.character(seq(18.5, 25, by = 5)))
p4

p5 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.3, 25, by = 5),
                       limits = c(18.3, 25),
                       labels = as.character(seq(18.3, 25, by = 5)))
p5
p6 <- ggplot(table, aes(x = logFC, y = -log10(p_adj.loc), color = AveExpr)) +
  geom_point() +
  labs(title = "RNAlistRDS005_df") +
  theme_minimal() +
  scale_colour_gradientn(colours = rev(rainbow(5)),
                       breaks = seq(18.5, 25, by = 5),
                       limits = c(18.5, 25),
                       labels = as.character(seq(18.5, 25, by = 5)))
p6
plot_grid(p1,p3,p2,p4,p6, ncol=3)

#Keep in mind to remove gene duplicates before giving ShSh the list
```

```{r df of table()s }
metadata <- read.csv("/home/common/data/output/projects/ENDO/E044/A006/Metadata_allE044_filtered.csv")
df <- data.frame(unclass(table(metadata$EndometriosisStatus,metadata$CyclePhases)))
df <- df %>%
  rename(Mid.Secretory = Secretory) %>%
  select(Proliferative, Early.Secretory, Mid.Secretory, Late.Secretory)

```

```{r VlnPlots of SLPI and NFKBIA for epithelial cells}
epithelial <- readRDS("/home/common/data/output/projects/ENDO/E044/A066/epithelial.rds")
#Filter by proliferative samples used for the DEG analysis
samples <- unique(read.csv("/home/common/data/output/projects/ENDO/E044/A030/major_celltypes_Tan/celltype_freq_sample.csv")$Var2)
Idents(epithelial) <- epithelial$sample
epithelial_prolif <- subset(epithelial, idents = samples)

#VlnPlot
p1 <- VlnPlot(epithelial_prolif, features = c("SLPI"), assay = "RNA",group.by = "sample", pt.size = 0,split.by = "EndometriosisStatus") + ylim(0,100) +
stat_summary(fun.y = median, geom='point', size = 5, colour = "black", shape = 95) + ggtitle("SLPI, RNA assay")
p2 <- VlnPlot(epithelial_prolif, features = c("NFKBIA"), assay = "RNA",group.by = "sample", pt.size = 0,split.by = "EndometriosisStatus") +
stat_summary(fun.y = median, geom='point', size = 5, colour = "black", shape = 95) + ggtitle("NFKBIA, RNA assay")
p3 <- VlnPlot(epithelial_prolif, features = c("SLPI"), assay = "SCT",group.by = "sample", pt.size = 0,split.by = "EndometriosisStatus") +
stat_summary(fun.y = median, geom='point', size = 5, colour = "black", shape = 95)+ ggtitle("SLPI, SCT assay")
p4 <- VlnPlot(epithelial_prolif, features = c("NFKBIA"), assay = "SCT",group.by = "sample", pt.size = 0,split.by = "EndometriosisStatus") +
stat_summary(fun.y = median, geom='point', size = 5, colour = "black", shape = 95)+ ggtitle("NFKBIA, SCT assay")
plot_grid(p1,p3,p2,p4, ncol=2)
ggsave(plot = plot_grid(p1,p3,p2,p4, ncol=2), paste0(dir_out, "SLPI_NFKBIA_epithelial_prolif_RNASCT.png"))

##Calculate mean/median SLPI/NKKBIA expression per sample for SCT assay
expression_values <- GetAssayData(object = epithelial_prolif, assay = "SCT")[c("SLPI", "NFKBIA"), ]
expression_df <- as.data.frame(t(as.matrix(expression_values)))
expression_df$sample <- gsub("_.*", "", rownames(expression_df))
mean_values <- aggregate(. ~ sample, data = expression_df, FUN = mean)
median_values <- aggregate(. ~ sample, data = expression_df, FUN = median)
# Combine mean and median values into a single table
result_table_SCT <- data.frame(
  sample_SCT = mean_values$sample,
  SLPI_mean_SCT = mean_values$SLPI,
  SLPI_median_SCT = median_values$SLPI,
  NFKBIA_mean_SCT = mean_values$NFKBIA,
  NFKBIA_median_SCT = median_values$NFKBIA
)

##Calculate mean/median SLPI/NKKBIA expression per sample for RNA assay
expression_values <- GetAssayData(object = epithelial_prolif, assay = "RNA")[c("SLPI", "NFKBIA"), ]
expression_df <- as.data.frame(t(as.matrix(expression_values)))
expression_df$sample <- gsub("_.*", "", rownames(expression_df))
mean_values <- aggregate(. ~ sample, data = expression_df, FUN = mean)
median_values <- aggregate(. ~ sample, data = expression_df, FUN = median)
# Combine mean and median values into a single table
result_table_RNA <- data.frame(
  sample = mean_values$sample,
  SLPI_mean_RNA = mean_values$SLPI,
  SLPI_median_RNA = median_values$SLPI,
  NFKBIA_mean_RNA = mean_values$NFKBIA,
  NFKBIA_median_RNA = median_values$NFKBIA
)

#combine RNA and SCT assay and add EndometriosisStatus condition to df
result_table <- cbind(result_table_RNA,result_table_SCT)
sample_EndometriosisStatus <- epithelial_prolif@meta.data %>% select(c("sample","EndometriosisStatus")) %>% distinct()
result_table <- left_join(result_table,sample_EndometriosisStatus, by = "sample")
write.csv(result_table,"mean_median_SLPI_NFKBIA_epithelial_prolif_RNASCT.csv")

```

```{r Make vulcano plots for DEGs colored by major cell types}
##RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv")
##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes
mt.genes <- grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#mitochondrial genes
rb.genes <- grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE)#ribosomal genes
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  dplyr::filter(!gene %in% c(hb.genes, mt.genes, rb.genes)) %>%
  dplyr::filter(annotation_reference == "Tan") %>%
  dplyr::filter(AveExpr > 18.53202)
##Rename cluster_id to major cell types
meta <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_metadata.rds")
meta$Symphony_Refined[meta$Symphony_Refined == "M$\\Phi$1-LYVE1"] <- "M-Phi-1-LYVE1"
meta$Symphony_Refined[meta$Symphony_Refined == "glandular early-secretory"] <- "glandular-early-secretory"
meta <- meta %>% 
  select(AnnotationMain,Symphony_Refined,Symphony_Refined_A029_United) %>%
  distinct()

RNAlistRDS005_df <- RNAlistRDS005_df %>%
  mutate(AnnotationMain = cluster_id)
RNAlistRDS005_df$AnnotationMain <- case_when(
  RNAlistRDS005_df$AnnotationMain %in% meta$Symphony_Refined ~ meta$AnnotationMain[match(RNAlistRDS005_df$AnnotationMain, meta$Symphony_Refined)],
  RNAlistRDS005_df$AnnotationMain %in% meta$Symphony_Refined_A029_United ~ meta$AnnotationMain[match(RNAlistRDS005_df$AnnotationMain, meta$Symphony_Refined_A029_United)],
  TRUE ~ RNAlistRDS005_df$AnnotationMain)
RNAlistRDS005_df$AnnotationMain[RNAlistRDS005_df$AnnotationMain == "myeloid" |
                                      RNAlistRDS005_df$AnnotationMain == "lymphocyte"] <- "immune"

##Plot
ggplot(RNAlistRDS005_df, aes(x = logFC, y = -log10(p_adj.loc), color = AnnotationMain)) +
  geom_point(size = 1) +
  #theme_minimal() +
  facet_wrap(~stage)
ggsave(filename = paste0(dir_out, 'vulcano_DEG_stage_main4.pdf'), width =7,height=4)

ggplot(RNAlistRDS005_df, aes(x = AnnotationMain, fill = cluster_id)) +
  geom_bar(stat = "count", position = "stack", width = 0.7) +  
  # theme_minimal() +  
  facet_wrap(~stage) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggsave(filename = paste0(dir_out, 'vulcano_DEG_stage_main5.pdf'), width =7,height=4)

```

```{r Captum scores}
#Load captum scores and combine tables
all_cells_corrgpca <- read.csv("/home/common/data/output/projects/ENDO/E042/A011/sigs/weighted_score_sig_corrgpca_eval_20pc_pcs.csv")
all_cells_corrgpca$dataset <- "all_cells_corrgpca"
all_cells_pca <- read.csv("/home/common/data/output/projects/ENDO/E042/A012/sigs/weighted_score_sig_pca_1pc.csv")
all_cells_pca$dataset <- "all_cells_pca"

##RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv")
##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes
mt.genes <- grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#mitochondrial genes
rb.genes <- grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE)#ribosomal genes
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  dplyr::filter(!gene %in% c(hb.genes, mt.genes, rb.genes)) %>%
  dplyr::filter(annotation_reference == "Tan") %>%
  dplyr::filter(AveExpr > 18.53202)

#Combine tables
all_cells <- rbind(all_cells_pca,all_cells_corrgpca)
all_cells <- left_join(all_cells, RNAlistRDS005_df, by = "gene") %>% 
  arrange(desc(rank))

```

```{r cycling fibroblasts}

#LTs annotation for cycling fibroblasts
data <- readRDS("/home/common/data/output/projects/ENDO/E031/A004/proliferative_SCT_percent.mt_sample_integration_batch_hvg_annotated.RDS")
meta <- data@meta.data
rm(data)
gc()
table(meta$Annotation, meta$EndometriosisStatus)

#Make cell id compatible with new data set
meta$cell_id <- paste(meta$sample, meta$cell.names, sep = '_')
rownames(meta) <- meta$cell_id

#Load entire dataset and add LTs annotation
data_ds <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")
data_ds <- AddMetaData(data_ds,meta[,c("Annotation", "scds.cxds.score")])
DimPlot(data_ds_new,group.by = "Annotation")

#Highlight cycling fibroblasts
Idents(data_ds) <- data_ds$Annotation
table(data_ds$Annotation)
cyclingfibs <- WhichCells(data_ds, idents = c("Cycling Fibroblasts"))
DimPlot(data_ds, reduction = "umap", cells.highlight = cyclingfibs, cols.highlight = c("darkred"), cols= "grey", raster = FALSE)

```

```{r Barplots for  filter response median rank E042A027}
#Load metadata
meta_final <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")

#Set variables to loop over
DIMRED <- c("corrgpca","pca")
ANNOTATION <- c("Symphony_Refined","AnnotationMain")
MEDIAN <- c("median_rank_endo_logic003","median_rank_control_logic003","median_rank_endo_logic005","median_rank_control_logic005","median_rank_endo_logic01","median_rank_control_logic01")

for (dim in DIMRED) {
#Read filter ranking
average_filter_ranking <- readRDS(paste0("/home/common/data/output/projects/ENDO/E042/A027/response_cellcnn_manually/all_cells/",dim,"_average_filter_ranking.rds"))

#Combine tables
meta <- left_join(meta_final, average_filter_ranking %>% select(cell_ident, median_rank_endo, median_rank_control), by = "cell_ident")

#Exclude NA
metaNAfree <- meta %>%
  select(median_rank_endo, median_rank_control,Symphony_Refined,AnnotationMain) %>%
  na.omit() %>%
  filter(Symphony_Refined != "granulocyte")

endo_threshold003 <- quantile(metaNAfree$median_rank_endo, 0.03)
control_threshold003 <- quantile(metaNAfree$median_rank_control, 0.03)
endo_threshold005 <- quantile(metaNAfree$median_rank_endo, 0.05)
control_threshold005 <- quantile(metaNAfree$median_rank_control, 0.05)
endo_threshold01 <- quantile(metaNAfree$median_rank_endo, 0.1)
control_threshold01 <- quantile(metaNAfree$median_rank_control, 0.1)

# Create the new column with ranking logic threshold
metaNAfree <- metaNAfree %>%
  mutate(median_rank_endo_logic003 = median_rank_endo <= endo_threshold003) %>%
  mutate(median_rank_control_logic003 = median_rank_control <= control_threshold003) %>%
  mutate(median_rank_endo_logic005 = median_rank_endo <= endo_threshold005) %>%
  mutate(median_rank_control_logic005 = median_rank_control <= control_threshold005) %>%
  mutate(median_rank_endo_logic01 = median_rank_endo <= endo_threshold01) %>%
  mutate(median_rank_control_logic01 = median_rank_control <= control_threshold01)

# Create bar plots
dir.create(paste0(dir_out,"median_rank_plots/"))

for (ann in ANNOTATION) {
  for (med in MEDIAN) {
    # Create plot
    plot1 <- ggplot(metaNAfree, aes_string(x = ann)) +
      geom_bar(position = "identity") +
      scale_y_continuous(trans = "log10") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
      ggtitle(paste0(dim, "___",ann, "___", med))
    plot2 <- ggplot(metaNAfree, aes_string(x = ann, fill = med)) +
      geom_bar(position = "fill") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
    combined_plot <- plot1 / plot2
    # Save plot
    ggsave(filename = paste0(dir_out, "/median_rank_plots/", dim, "___",ann, "___", med, ".pdf"), plot = combined_plot, width = 13, height = 10)
  }
}
}

```

```{r UMAPs for  filter response median rank E042A027}
#Load data and metadata
data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033.rds")
#data <- readRDS("/home/common/data/output/projects/ENDO/E044/A033/dataA003A033_downsampled.rds")
meta_final <- readRDS("/home/common/data/output/projects/ENDO/E044/A074/dataA003A033_metadata_ptime_snn_Final.rds")
average_filter_ranking <- readRDS(paste0("/home/common/data/output/projects/ENDO/E042/A027/response_cellcnn_manually/all_cells/pca_average_filter_ranking.rds"))
#Combine meta, clean up and do ranking
meta <- left_join(meta_final, average_filter_ranking %>% select(cell_ident, median_rank_endo, median_rank_control), by = "cell_ident") 
#Exclude NA
metaNAfree <- meta %>%
  select(median_rank_endo, median_rank_control,Symphony_Refined,AnnotationMain,cell_ident) %>%
  na.omit()
#Ranking
endo_threshold01 <- quantile(metaNAfree$median_rank_endo, 0.1)
control_threshold01 <- quantile(metaNAfree$median_rank_control, 0.1)
metaNAfree <- metaNAfree %>%
  mutate(median_rank_endo_logic01 = median_rank_endo <= endo_threshold01) %>%
  mutate(median_rank_control_logic01 = median_rank_control <= control_threshold01) %>%
  select(Symphony_Refined, median_rank_endo_logic01, median_rank_control_logic01,cell_ident)
#Make summary table
summary_table <- metaNAfree %>%
  select(Symphony_Refined, median_rank_endo_logic01, median_rank_control_logic01) %>%
  group_by(Symphony_Refined) %>%
  summarize(
    percent_endo_TRUE = mean(median_rank_endo_logic01) * 100,
    percent_control_TRUE = mean(median_rank_control_logic01) * 100
  )
write.csv(summary_table, paste0(dir_out,"summary_pca_average_filter_ranking_10perc.csv"))

#Highlight 10% cells
endo_true_cells <- subset(metaNAfree, median_rank_endo_logic01 == TRUE)$cell_ident
control_true_cells <- subset(metaNAfree, median_rank_control_logic01 == TRUE)$cell_ident
p1 <- DimPlot(data, reduction = "umap", cells.highlight = endo_true_cells, cols.highlight = c("darkred"), cols= "grey", raster = FALSE) + ggtitle("endo_true_cells")
p2 <- DimPlot(data, reduction = "umap", cells.highlight = control_true_cells, cols.highlight = c("darkred"), cols= "grey", raster = FALSE) + ggtitle("control_true_cells")

```

```{r make DEG table for paper}
#RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv") %>%
  filter(annotation_reference == "Tan") %>%
  filter(AveExpr > 18.53202)
##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes, there are no hb genes in the DEG list
RNAlistRDS005_df[RNAlistRDS005_df$gene %in% hb.genes, ]#no hb genes
#mitochondrial genes
mt.genes <- grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#mitochondrial genes, no genes
rb.genes <- grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE)#ribosomal genes, 77 genes
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  dplyr::filter(!gene %in% rb.genes)

write.csv(RNAlistRDS005_df, paste0(dir_out,"RNAlistRDS005_AveExpr18.53202_noRBgenes.csv"))

length(unique(RNAlistRDS005_df$gene)) #427
```

```{r}

metadata <- read.csv('/home/common/data/output/projects/ENDO/E042/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv')
metadata$new_column <- NULL
metadata$batch <- NULL

# Remove 'names' column as it will be used as row labels
row_names <- metadata$names
metadata <- metadata[, -1]

# Convert non-numeric columns to factors
metadata[, c("EndometriosisStatus", "lane",  "endo_EndometriosisGrade",  "fold1", "fold2", "fold3", "fold4", "fold5")] <- lapply(metadata[, c("EndometriosisStatus", "lane", "endo_EndometriosisGrade", "fold1", "fold2", "fold3", "fold4", "fold5")], as.factor)

# Create heatmap using ComplexHeatmap
pdf(paste0(dir_out, "metadata_prolif_eval_fold_all_heatmap.pdf"), width = 7, height = 9)
Heatmap(metadata,
        name = "Metadata",
        row_names_side = "left",
        row_names_gp = gpar(fontsize = 8),
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        show_column_names = TRUE,
        show_row_names = TRUE,
        column_title = "Columns",
        row_title = "Row Names",
        heatmap_legend_param = list(title = "Legend")
)
dev.off()

######################
metadata <- read.csv('/home/common/data/output/projects/ENDO/E042/A001/summary_plots_cv_splits/metadata_prolif_eval_fold_all.csv')
metadata$new_column <- NULL
metadata$batch <- NULL

# Remove 'names' column as it will be used as row labels
row_names <- metadata$names
metadata <- metadata[, -1]

# Convert non-numeric columns to factors
metadata[, c("EndometriosisStatus", "lane","endo_EndometriosisGrade",  "fold1", "fold2", "fold3", "fold4", "fold5")] <- lapply(metadata[, c("EndometriosisStatus", "lane", "endo_EndometriosisGrade", "fold1", "fold2", "fold3", "fold4", "fold5")], as.factor)

# Convert remaining columns to a matrix
mat <- data.matrix(metadata)

# Create heatmap
pdf(paste0(dir_out, "metadata_prolif_eval_fold_all_heatmap2.pdf"), width = 7, height = 9)
heatmap.2(mat,
          Rowv=FALSE, Colv=FALSE,
          dendrogram="none",
          trace="none",
          col = heat.colors(256),
          margins=c(8,10),
          labRow=row_names,
          scale="column",
          key=TRUE,
          keysize=1.5,
          key.title="Color Key",
          key.xlab="Value",
          key.ylab=NULL,
          density.info="none",
          main="Heatmap of Metadata")
dev.off()

```