### 2025-02-23 Figures/Preprocessing_metadata.Rmd
Complemented Figure script

```{r setup}
library(Seurat)
library(dplyr)
library(tibble)
library(stringr)
library(ggplot2)
library(cowplot)
library(ggbreak) 
library(tidyr)
library(patchwork)
library(ComplexUpset)
library(ggrepel)
library(Polychrome)
library(ComplexHeatmap)
library(gplots)
library(RColorBrewer)
library(randomcoloR)

```

ln -s /home/common/data/output/projects/ENDO/E044/A033/EndoAtlas_GEO_meta_complete_downsampled.rds EndoAtlas.rds
EndoAtlas_meta <- readRDS(paste0(dir_data,"EndoAtlas_meta.rds"))
EndoAtlas <- readRDS(paste0(dir_data,"EndoAtlas.rds"))
EndoAtlas <- AddMetaData(EndoAtlas, EndoAtlas_meta)
saveRDS(EndoAtlas,paste0(dir_data,"EndoAtlas.rds") )

Add additional data to the Seurat metadata and reformat
```{r plotting of cell types per sample}
#Set directory links and create directories
dir_ENDO <- "/home/duempelmann/analysis-projects/ENDO/E044/A085/endometriosis_endometrium_scRNA_atlas/"

dir_data <- paste0(dir_ENDO,"_Data/")
dir_out <- paste0(dir_ENDO,"Figures_out/")
if (!dir.exists(dir_out)) {
  dir.create(dir_out, recursive = TRUE)
}

#Load data
EndoAtlas <- readRDS(paste0(dir_data, "EndoAtlas.rds"))
EndoAtlas_meta <- EndoAtlas@meta.data

#Preserve cell names
EndoAtlas_meta$cell_names <- rownames(EndoAtlas_meta)

##Modify menstrual cycle information
# Create column MenstrualCyclePhase_main based on MenstrualCyclePhase by merging the secretory sub-phases from 
EndoAtlas_meta$MenstrualCyclePhase_main <- as.character(EndoAtlas_meta$MenstrualCyclePhase)
EndoAtlas_meta$MenstrualCyclePhase_main[EndoAtlas_meta$MenstrualCyclePhase_main %in% c("secretory_early", "secretory_mid", "secretory_late")] <- "secretory"
#Factor samples by MenstrualCyclePhase_main for plotting
EndoAtlas_meta$MenstrualCyclePhase_main <- factor(EndoAtlas_meta$MenstrualCyclePhase_main, levels = c("proliferative","periovulatory","secretory"))

##Add additional metadata from manuscript Supplementary Table 1
# Read the tab-delimited text file
SupplementaryTable1 <- read.delim(paste0(dir_data, "SupplementaryTable1.txt"), header = TRUE, stringsAsFactors = FALSE)
# Perform a left join of the `meta` dataframe with `SupplementaryTable1`
EndoAtlas_meta <- dplyr::left_join(
  EndoAtlas_meta,
  SupplementaryTable1 %>% 
    dplyr::select(
      Patient.sample.ID,
      epithelial.pseudotime,
      sequencing.batch,
      X10x.capturing.batch,
      Minor.exclusion.criteria.logical,
      Menstrual.cycle.phase.marker.analysis,
      Cell.type.frequency.analysis..Figure.2e.f.,
      DEG.analysis..Figure.3a.b.,
      Ligand.receptor.analysis..Figure.3c.h.
    ),
  by = c("sample" = "Patient.sample.ID")
)

##Determine order of samples by MenstrualCyclePhase and epithelial.pseudotime
MenstrualCyclePhase_sampleOrder <- EndoAtlas_meta%>%
  select(sample, MenstrualCyclePhase,epithelial.pseudotime) %>%
  distinct() %>%
  arrange(MenstrualCyclePhase, epithelial.pseudotime) %>%
  mutate(MenstrualCyclePhase_sampleOrder = row_number()) %>%
  ungroup()
#Factor samples by MenstrualCyclePhase and epithelial.pseudotime for plotting
EndoAtlas_meta$sample <- factor(EndoAtlas_meta$sample, levels = c(MenstrualCyclePhase_sampleOrder$sample))

#Add cell names as row names
rownames(EndoAtlas_meta) <- EndoAtlas_meta$cell_names

#Save EndoAtlas_meta as .rds
saveRDS(EndoAtlas_meta, paste0(dir_data, "EndoAtlas_meta.rds"))

#EndoAtlas_meta <- readRDS(paste0(dir_data, "EndoAtlas_meta.rds"))

#Save color pals
qual_col_pals <- brewer.pal.info[brewer.pal.info$category == "qual", ]
col_vector <- unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```

Save epithelial object and upload to GEO.
```{r}
#Load monocl object
data_monocl <- readRDS("/home/common/data/output/projects/ENDO/E044/A062/trajectory_day5-6-7_greenCircle/trajectory_timepoint0.rds")
listData <- (data_monocl@colData@listData)
listData <- listData["cell_ident"]

EndoAtlas_meta <- readRDS(paste0(dir_data, "EndoAtlas_meta.rds"))
colnames(EndoAtlas_meta)

########################
# Extract cell.ident from listData
cell_ids <- listData[["cell_ident"]]

# Ensure cell_ids is not empty
if (length(cell_ids) == 0) {
  stop("cell.ident in listData is empty. Ensure it is populated before proceeding.")
}

# Filter and order EndoAtlas_meta by cell_ids
filtered_meta <- EndoAtlas_meta[EndoAtlas_meta$cell_names %in% cell_ids, ]
ordered_meta <- filtered_meta[match(cell_ids, filtered_meta$cell_names), ]

# Verify alignment
if (!all(ordered_meta$cell_names == cell_ids)) {
  stop("Ordering of filtered EndoAtlas_meta does not match cell_ids.")
}

# Add columns from ordered_meta to listData
columns_to_add <- c("sample", "EndometriosisStatus", "EndometriosisGrade", "CycleDay", 
                    "ProgesteroneSerum", "MenstrualCyclePhase", "AnnotationMain", 
                    "AnnotationRefined", "cell_names", "MenstrualCyclePhase_main")

for (col in columns_to_add) {
  if (col %in% colnames(ordered_meta)) {
    listData[[col]] <- ordered_meta[[col]]
  } else {
    warning(paste("Column", col, "not found in EndoAtlas_meta. Skipping."))
  }
}

# Update the main object
data_monocl@colData@listData <- listData

data_monocl@principal_graph_aux@listData[["UMAP"]][["pseudotime"]]

saveRDS(data_monocl, paste0(dir_data, "epithelial_cells_monocle3.rds"))
```